---
df_print: tibble
output:
  html_notebook:
    code_folding: hide
  pdf_document:
    keep_tex: yes
always_allow_html: yes
---

```{r misc-setup, echo = FALSE, warning=FALSE,message=FALSE,comment=FALSE}
library(magrittr)
library(operator.tools)
library(knitr)
library(rprojroot)
library(tidyverse)
library(rgdal)
library(sp)
library(rgeos)
library(miscgis)
library(tigris)
library(leaflet)
library(ggthemes)
library(stringr)
library(downloader)
library(miscgis)
library(sf)
root <- rprojroot::is_rstudio_project
root_file <- root$make_fix_file()
opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, comment=FALSE)
green <- miscgis::miscgis_pals$tableau_cat[["green"]]
blue <- miscgis::miscgis_pals$tableau_cat[["blue"]]
orange <- miscgis::miscgis_pals$tableau_cat[["orange"]]
red <- miscgis::miscgis_pals$tableau_cat[["red"]]
teal <- miscgis::miscgis_pals$tableau_cat[["teal"]]
pal_rgb_4 <- miscgis_pals$tableau_cat[c("red","gold","green","blue")] %>% unlist %>% palette()
```

###Seattle Boundary

```{r misc-sea-bound, fig.cap="Seattle\'s geographic boundary"}

if(!file.exists(root_file('1-data/4-interim/seattle-sf.rds'))){
        tigris::places(state = "WA") %>%
                tigris::filter_place(place = "Seattle") %>%
                spTransform(CRSobj = crs_proj) %>% 
                st_as_sf() %>% 
                miscgis::coerce_to_geom(sf::st_multipolygon) %>% 
                write_rds(root_file('1-data/4-interim/seattle-sf.rds'))
        
}

sea_sf <- read_rds(root_file('1-data/4-interim/seattle-sf.rds'))

green <- miscgis::miscgis_pals$tableau_cat[["green"]]

myLfltGrey(data = as(sea_sf,'Spatial')) %>% addPolygons(color = green,opacity = 1,fillColor = green,fillOpacity = .5)

```

###King County Subdivision Boundary

```{r misc-sea-ccd, fig.cap="Seattle Subdivision of King County\'s geographic boundary"}

if(!file.exists(root_file('1-data/4-interim/seattle-ccd-sf.rds'))){
        tigris::county_subdivisions(state = "53",county = "033") %>% 
                subset(NAME == "Seattle") %>% 
                spTransform(CRSobj = crs_proj) %>% 
                st_as_sf() %>% 
                miscgis::coerce_to_geom(sf::st_multipolygon) %>% 
                write_rds(root_file('1-data/4-interim/seattle-ccd-sf.rds'))
}

sea_ccd_sf <- read_rds(root_file('1-data/4-interim/seattle-ccd-sf.rds'))

orange <- miscgis::miscgis_pals$tableau_cat[["orange"]]

myLfltGrey(data = as(sea_ccd_sf,'Spatial')) %>% addPolygons(color = orange,opacity = 1,fillColor = orange,fillOpacity = .5)

```

### Tracts in Seattle CCD

In the absence of a straight-forward method for identifying all the census tracts in the Seattle CCD subdivision of King County, it is possible to extract this information from [American Factfinder](https://factfinder.census.gov/faces/nav/jsf/pages/index.xhtml). This [tutorial](https://ask.census.gov/faq.php?id=5000&faqId=1605) describes how to use the American Factfinder interface to extract a list of all "all tracts within (or partially within) a census place"; substituting "county subdivision" for "place" will retrieve the desired results.

```{r misc-tracts-ccd}


if(!file.exists(root_file('1-data/4-interim/tr-ccd-sf.rds'))){
        tr_ccd_geoid <- 
                read_csv(
                        root_file('1-data/3-external/manual/seattle-ccd/ACS_12_5YR_B01001/ACS_12_5YR_B01001_with_ann.csv'), 
                        col_types = cols(Id2 = col_character()), 
                        skip = 1) %>% 
                mutate(NEW_GEOID1 = str_sub(Id2,1,5),
                       NEW_GEOID2 = str_sub(Id2,16,21),
                       GEOID = paste0(NEW_GEOID1,NEW_GEOID2)) %>% 
                select(GEOID) %>% unlist(use.names = F)
        
        # If the WA census tracts objects isn't yet loaded in the global env.,
        # download and load it
        if(!exists('tr')){
                tr <- 
                        tigris::tracts(state = "WA",county = "King", year = 2014) %>%
                        spTransform(CRSobj = crs_proj) %>% 
                        st_as_sf() %>% 
                        miscgis::coerce_to_geom(sf::st_multipolygon)
                }
        
subset(tr, GEOID %in% tr_ccd_geoid) %>% 
        st_as_sf %>% 
        miscgis::coerce_to_geom(st_multipolygon) %>% 
        write_rds(root_file('1-data/4-interim/tr-ccd-sf.rds'))
}

tr_ccd_sf <- read_rds(root_file('1-data/4-interim/tr-ccd-sf.rds'))

show_tr_ccd_sf <- function(){
        myLfltGrey() %>% 
        addPolygons(data = as(tr_ccd_sf,"Spatial"),weight = .5,color = blue,opacity = 1,fillColor = blue,fillOpacity = .5) %>% 
        addPolygons(data = as(sea_ccd_sf,"Spatial"),color = orange,opacity = 1,fillOpacity = 0)
}
# show_tr_ccd_sf()
```

###Puget Sound Waterbodies

These are useful for "clipping" census geographies whose boundaries extend into waterbodies.

```{r misc-wtr, fig.cap="Puget Sound waterbodies"}

if(!file.exists(root_file('1-data/4-interim/wtr_sf.rds'))){
        fp_wtr <- root_file('1-data/3-external/NHDMajor.gdb')

# check if the file already exists, if not then download it
if(!file.exists(fp_wtr)){
        
        url <- "ftp://www.ecy.wa.gov/gis_a/inlandWaters/NHD/NHDmajor.gdb.zip" # save the URL for the waterbodies data
        
        temp <- tempfile() # create a temporary file to hold the compressed download
        
        download(url, dest = temp, mode="wb") # download the file
        
        unzip (temp, exdir = root_file('1-data/3-external/')) # extract the ESRI geodatabase file to a project folder
}

wtr_sp <-
        suppressWarnings(readOGR(dsn = fp_wtr,      # create a waterbodies shape
                layer = "NHD_MajorWaterbodies",verbose = FALSE,pointDropZ = TRUE)) %>%
        gBuffer(byid=TRUE, width=0) %>% # clean up self-intersecting polygons
        spTransform(CRSobj = crs_proj)  # transform the projection to match the project projection
wtr_sf <-  
        wtr_sp %>% 
        st_as_sf() %>% 
        miscgis::coerce_to_geom(st_multipolygon)

wtr_sf %>% 
        st_intersects(x = sea_ccd_sf,y = .) %>% 
        unlist(use.names = F) %>% 
        wtr_sf[.,] %>% 
        write_rds(root_file('1-data/4-interim/wtr_sf.rds'))
}

wtr_sf <- read_rds(root_file('1-data/4-interim/wtr_sf.rds'))

show_wtr <- function(){
        myLfltGrey(data = as(wtr_sf,'Spatial')) %>% 
                addPolygons(color = blue, opacity = 1, 
                            weight = .5, fillColor = blue,fillOpacity = .5)        
}

show_wtr()

```
