---
df_print: tibble
output:
  html_notebook:
    default
  pdf_document:
    keep_tex: yes
always_allow_html: yes
---

```{r census-geos-setup, echo = FALSE, warning=FALSE,message=FALSE,comment=FALSE}
library(plyr)
library(knitr)
library(rprojroot)
library(rgdal)
library(sp)
library(rgeos)
library(tigris)
library(leaflet)
library(ggthemes)
library(magrittr)
library(stringr)
library(downloader)
library(webshot)
library(htmltools)
library(gplots)
library(ggmap)
library(shiny)
library(htmlwidgets)
library(readxl)
library(acs)
library(RColorBrewer)
library(tidyverse)
library(miscgis)
library(operator.tools)
library(sf)
root <- rprojroot::is_rstudio_project
root_file <- root$make_fix_file()
opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, comment=FALSE)

```

```{r census-geos-colors}
green <- miscgis::miscgis_pals$tableau_cat[["green"]]
blue <- miscgis::miscgis_pals$tableau_cat[["blue"]]
orange <- miscgis::miscgis_pals$tableau_cat[["orange"]]
red <- miscgis::miscgis_pals$tableau_cat[["red"]]
teal <- miscgis::miscgis_pals$tableau_cat[["teal"]]
pal_rgb_4 <- miscgis::miscgis_pals$tableau_cat[c("red","gold","green","blue")] %>% unlist %>% palette()
pal_rgb_4 <- miscgis::miscgis_pals$tableau_cat[c("red","gold","green","blue")] %>% unlist %>% palette()
pal_rgb_6 <- miscgis::miscgis_pals$tableau_cat[c("red","gold","green","blue","orange","purple")] %>% unlist %>% palette()
pal_rgb_6 <- miscgis::miscgis_pals$tableau_cat[c("red","gold","green","blue","orange","purple")] %>% unlist %>% palette()
```

### Introduction
The primary geographic unit of this assessment is the [census tract](https://www.census.gov/geo/reference/gtc/gtc_ct.html). As is the case with many communities, the census geographies do not coincide exactly with the formal geographic boundary of the study's three sites, and should be considered as _spatial approximations_ of these communities.

Listed below are the geographic identifiers of the census tracts that approximate each site.


```{r census-geos-intro, results="asis"}

# Rainier Valley
rv <- c("53033010001", "53033010300", "53033010401", "53033011001", "53033011002", "53033011101","53033011102", "53033011700", "53033011800", "53033011900")

# White Center
wc <- c("53033026600","53033026700","53033026500","53033026801","53033026802","53033027000")

# SeaTac/Tukwila

stc_hus <- read_csv(root_file('1-data/3-external/manual/seatac-tr/DEC_10_SF1_H1/DEC_10_SF1_H1_with_ann.csv'), 
                    col_types = cols(Id2 = col_character()), skip = 1)

tuk_hus <- read_csv(root_file('1-data/3-external/manual/tukwila-tr/DEC_10_SF1_H1/DEC_10_SF1_H1_with_ann.csv'), 
    col_types = cols(Id2 = col_character()), 
    skip = 1)

stc <-  paste(substr(x = stc_hus$Id2,1,2),substr(x = stc_hus$Id2,8,16),sep = "")

tuk <-  paste(substr(x = tuk_hus$Id2,1,2),substr(x = tuk_hus$Id2,8,16),sep = "")

stctuk <- c(stc,tuk)

coo <- c(rv,wc,stctuk)

df <- miscgis::cbind_fill(rv,wc,stctuk) %>% 
        as_tibble() %>% 
        `colnames<-`(c("Rainier Valley",
                       "White Center",
                       "SeaTac/Tukwila"))

knitr::kable(df,caption = "Census Tract GEOIDs")
```

```{r census-geos-ycc-nhoods}

if(!file.exists(root_file('1-data/4-interim/coo-geos-sf.rds'))){
        
        tr_ccd_coo_sf <- 
                read_rds(root_file('1-data/4-interim/tr-ccd-sf.rds')) %>% 
                mutate(NAME_FULL = case_when(.$GEOID %in% rv ~ 'Rainier Valley',
                                             .$GEOID %in% wc ~ 'White Center',
                                             .$GEOID %in% stctuk ~ 'SeaTac/Tukwila',
                                             .$GEOID %!in% coo ~ NA_character_),
                       NAME = case_when(.$GEOID %in% rv ~ 'RV',
                                             .$GEOID %in% wc ~ 'WC',
                                             .$GEOID %in% stctuk ~ 'STKW',
                                             .$GEOID %!in% coo ~ NA_character_)) %>%
                select(NAME,NAME_FULL,GEOID,GEOID6 = TRACTCE,geometry)
        
        
        # Tract-communities:
        # dissolve YCC tracts by Community, join to the tract data, 
        # record the geography type in a variable to make it easy
        # to split the dataset later.
        
        tr_ccd_coo_sf <- 
                tr_ccd_coo_sf %>% 
                filter(!is.na(NAME)) %>% 
                group_by(NAME) %>% 
                summarise_all(first) %>% 
                mutate(GEOID = NA_character_,
                       GEOID6 = NA_character_) %>% 
                rbind(tr_ccd_coo_sf) %>% 
                mutate(GEOGRAPHY = ifelse(is.na(GEOID),
                                          'community',
                                          'tract')) %>% 
                miscgis::coerce_to_geom(st_multipolygon) %>% 
                select(everything(),geometry) %>% 
                st_set_crs(st_crs(tr_ccd_coo_sf))
        
        # COO Communities (Combined)
        
        coo_sf <- 
                tr_ccd_coo_sf %>% 
                filter(GEOGRAPHY == 'community') %>% 
                mutate(GRP = 1) %>% 
                group_by(GRP) %>% 
                summarise_all(first) %>% 
                select(-GRP) %>% 
                mutate(NAME = 'COO',
                       NAME_FULL = 'COO Communities (combined)',
                       GEOGRAPHY = 'communities (multiple)') %>% 
                miscgis::coerce_to_geom(st_multipolygon)
        
        # Seattle CCD
        
        sea_ccd_sf <-  
                tr_ccd_coo_sf %>% 
                filter(GEOGRAPHY == 'tract') %>% 
                mutate(GRP = 1) %>% 
                group_by(GRP) %>% 
                summarise_all(first) %>% 
                select(-GRP) %>% 
                mutate(NAME = 'SEACCD',
                       NAME_FULL = 'Seattle CCD',
                       GEOGRAPHY = 'county subdivision',
                       GEOID = NA_character_,
                       GEOID6 = NA_character_) %>% 
                miscgis::coerce_to_geom(st_multipolygon)
        
        
        coo_geos_sf <- 
                rbind(tr_ccd_coo_sf,coo_sf) %>% 
                rbind(sea_ccd_sf) %>% 
                mutate(JOIN_ID = ifelse(is.na(GEOID6),NAME,GEOID6),
                       NAME_FCT = factor(NAME,
                                              levels = c('RV',
                                                         'WC',
                                                         'STKW',
                                                         'COO',
                                                         'SEACCD'),
                                              ordered = TRUE),
                       NAME_FULL_FCT = factor(NAME_FULL,
                                              levels = c('Rainier Valley',
                                                         'White Center',
                                                         'SeaTac/Tukwila',
                                                         'COO Communities (combined)',
                                                         'Seattle CCD'),
                                              ordered = TRUE),
                       NAME_FULL_FCT_COMMS = factor(NAME_FULL,
                                 levels = c('Rainier Valley',
                                                         'White Center',
                                                         'SeaTac/Tukwila'),
                                 ordered = TRUE)) %>% 
                st_set_crs(st_crs(tr_ccd_coo_sf))
        
        coo_geos_sf %>%
                write_rds(root_file('1-data/4-interim/coo-geos-sf.rds'))
                
        
}

# Read in the data

coo_geos_sf <- 
        read_rds(root_file('1-data/4-interim/coo-geos-sf.rds'))

```

```{r census-geos-ycc-summary, echo=TRUE}

# Print a summary (first 4 records of each geography type)
coo_geos_sf %>% 
        group_by(GEOGRAPHY) %>% 
        slice(1:3) %>% 
        arrange(GEOGRAPHY,NAME) 
```

### Maps
#### Seattle CCD Tracts
```{r census-geos-tr-map}

show_tr_ccd <- function(){
       
        coo_geos_sf %>% 
                filter(GEOGRAPHY %in% 'tract') %>% 
                as("Spatial") %>% 
                myLfltGrey() %>% 
                addPolygons() 
}

# show_tr_ccd()


```

#### COO Tract-Communities
```{r census-geos-coo-map, fig.height=1.25, fig.width=1,  fig.show='hold', dpi=150,fig.cap='Rainier Valley, White Center, and SeaTac/Tukwila'}

# Map Center

cntr <- coo_geos_sf %>% 
        filter(GEOGRAPHY %in% 'communities (multiple)') %>% 
        st_centroid() %>% st_bbox()
lng <- cntr[['xmin']]
lat <- cntr[['ymin']]
        
        
# Individiaul Maps


rv_sf <- 
        coo_geos_sf %>% 
        filter(GEOGRAPHY %in% 'tract' & NAME %in% 'RV')

wc_sf <- 
        coo_geos_sf %>% 
        filter(GEOGRAPHY %in% 'tract' & NAME %in% 'WC')

stkw_sf <- 
        coo_geos_sf %>% 
        filter(GEOGRAPHY %in% 'tract' & NAME %in% 'STKW')


# Rainier Valley
        
rv_map <- 
        rv_sf %>% 
        as("Spatial") %>% 
        miscgis::myLfltGrey() %>%
        addPolygons(smoothFactor = 0,
                    color = col2hex("white"),
                    weight = .5,
                    opacity = 1,
                    fillColor = blue,
                    fillOpacity = .65
        ) %>% 
        setView(lng,lat,zoom = 9) %>% 
        miscgis::styleWidget(style = "float:left;margin:1px")



# White Center

wc_map <- 
        wc_sf %>% 
        as("Spatial") %>% 
        miscgis::myLfltGrey() %>%
        addPolygons(smoothFactor = 0,
                    color = col2hex("white"),
                    weight = .5,
                    opacity = 1,
                    fillColor = green,
                    fillOpacity = .65
        ) %>% 
        setView(lng,lat,zoom = 9) %>% 
        miscgis::styleWidget(style = "float:left;margin:1px")

# SeaTac/Tukwila

stkw_map <- 
        stkw_sf %>% 
        as("Spatial") %>% 
        miscgis::myLfltGrey() %>%
        addPolygons(smoothFactor = 0,
                    color = col2hex("white"),
                    weight = .5,
                    opacity = 1,
                    fillColor = red,
                    fillOpacity = .65
        ) %>% 
        setView(lng,lat,zoom = 9) %>% 
        miscgis::styleWidget(style = "float:left;margin:1px")

rv_map

wc_map

stkw_map
       

```

