---
df_print: tibble
output:
  html_notebook:
    code_folding: hide
  pdf_document:
    keep_tex: yes
always_allow_html: yes
---


```{r demo-change-intro-setup, echo = FALSE, warning=FALSE,message=FALSE,comment=FALSE}
library(plyr)
library(knitr)
library(rprojroot)
library(rgdal)
library(sp)
library(rgeos)
library(tigris)
library(leaflet)
library(ggthemes)
library(magrittr)
library(stringr)
library(downloader)
library(webshot)
library(htmltools)
library(gplots)
library(ggmap)
library(shiny)
library(htmlwidgets)
library(readxl)
library(acs)
library(RColorBrewer)
library(tidyverse)
library(miscgis)
library(operator.tools)
library(leaflet.extras)
library(viridisLite)
library(sf)
root <- rprojroot::is_rstudio_project
root_file <- root$make_fix_file()
opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, comment=FALSE)
```

```{r demo-change-intro-load-objs}

# Load census indicators as an sf object
if(!exists('census_inds_sf')){census_inds_sf <- read_rds(root_file('./1-data/5-tidy/coo-acs-inds-not-acs-sf.rds'))}

inds_seaccd <- 
        crossing(c('VULN_PCT_2015_SEACCD','DEMO_CHNG_PCT_SEACCD'),
                c('B03002_EST','B15002_EST', 'B19001_EST', 'B25033_EST')) %>% 
        invoke_rows(str_c,.d = .,sep = '_',.collate = 'cols',.to = 'IND') %>% 
        extract2('IND') %>% 
        tibble('IND' = .,
               'TOPIC' = c('race', 'housing', 'ed', 'income') %>% rep(2),
               'CATEGORY' = c(rep('DEMO_CHNG',4),rep('VULN',4))) %>% 
        mutate(VALUE = map_dbl(IND, ~ census_inds_sf %>% unclass %>% as_tibble %>% select(matches(.x)) %>% slice(1) %>% extract2(1)),
               VALUE = scales::percent(VALUE),
               NAME = str_c(TOPIC,CATEGORY,sep = '_')) %>% 
        {set_names(x = .$VALUE, nm = .$NAME)}
```

A tract is considered to have experienced gentrification-related "demographic change" during the observation period (2004-2015) if any **two of the following four** indicators exceed the values observed in the county subdivision:

  * The white population share either increased or it decreased less than **`r inds_seaccd[['race_DEMO_CHNG']]`**
  * The share of homeowners either increased more than **`r inds_seaccd[['housing_DEMO_CHNG']]`**
  * The share of the population 25 years and older with a bachelor’s degree increased more
than **`r inds_seaccd[['ed_DEMO_CHNG']]`**
  * The share of household earning more than 80% AMI increased or it decreased less than **`r inds_seaccd[['income_DEMO_CHNG']]`**
  
 <br>