---
df_print: tibble
output:
  html_notebook:
    code_folding: hide
  pdf_document:
    keep_tex: yes
always_allow_html: yes
---

```{r chng-setup, echo = FALSE, warning=FALSE,message=FALSE,comment=FALSE}
library(plyr)
library(knitr)
library(rprojroot)
library(rgdal)
library(sp)
library(rgeos)
library(tigris)
library(leaflet)
library(ggthemes)
library(magrittr)
library(stringr)
library(downloader)
library(webshot)
library(htmltools)
library(gplots)
library(ggmap)
library(shiny)
library(htmlwidgets)
library(readxl)
library(acs)
library(RColorBrewer)
library(tidyverse)
library(miscgis)
library(operator.tools)
library(leaflet.extras)
library(viridisLite)
library(sf)
root <- rprojroot::is_rstudio_project
root_file <- root$make_fix_file()
opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, comment=FALSE)
```

### Introduction

```{r chng-colors}
green <- miscgis::miscgis_pals$tableau_cat[["green"]]
blue <- miscgis::miscgis_pals$tableau_cat[["blue"]]
orange <- miscgis::miscgis_pals$tableau_cat[["orange"]]
red <- miscgis::miscgis_pals$tableau_cat[["red"]]
teal <- miscgis::miscgis_pals$tableau_cat[["teal"]]
pal_rgb_4 <- miscgis::miscgis_pals$tableau_cat[c("red","gold","green","blue")] %>% unlist %>% palette()
pal_rgb_4 <- miscgis::miscgis_pals$tableau_cat[c("red","gold","green","blue")] %>% unlist %>% palette()
pal_rgb_6 <- miscgis::miscgis_pals$tableau_cat[c("red","gold","green","blue","orange","purple")] %>% unlist %>% palette()
pal_rgb_6 <- miscgis::miscgis_pals$tableau_cat[c("red","gold","green","blue","orange","purple")] %>% unlist %>% palette()
proj_light_grey <- col2hex("grey75")
proj_grey <- col2hex("grey50")
proj_dark_grey <- col2hex("grey25")
```

_NOTE: this text is placeholder and must be changed_

We defined census tracts with gentrification-related demographic change from 2005-2009 to 2010-2014 as
those that experienced either at least 3 of the following 4:

  * The white population share either increased or it decreased less than X.0%
  * The share of homeowners either increased or it decreased less than X%
  * The share of the population 25 years and older with a bachelor’s degree increased more
than X%
  * The share of household earning more than 80% AMI increased or it decreased less than X%
or experienced only 2 out of 4, which were:
  * The white population share either increased or it decreased less than X%
  * The share of the population 25 years and older with a bachelor’s degree increased more
than X%

### Change in Share of Home Owners (%), 2005-2009 to 2010-2014

#### Plot
```{r chng-own-2014-acs}

# Download the Census table for the project's geographies, 2010-2014
# -----------------------------------------------------------------------------

coo_geo_acs <- read_rds(root_file('1-data/4-interim/coo-geo-acs.rds'))

chng_own_tbl <- "B25033"  # census table code

# 2010-2014, 5-year estimates
if(!file.exists(root_file('1-data/4-interim/chng-own-2014-orig-acs.rds'))){
        
        acs.fetch(endyear = 2014, geography = coo_geo_acs, 
                  table.number = chng_own_tbl) %>%
                write_rds(root_file('1-data/4-interim/chng-own-2014-orig-acs.rds'))
        
}

chng_own_2014_orig_acs <- read_rds(root_file('1-data/4-interim/chng-own-2014-orig-acs.rds'))


# Process and save the data
# -----------------------------------------------------------------------------

if(!file.exists(root_file('1-data/4-interim/chng-own-2014-acs.rds'))){
        chng_own_2014_acs_1 <- chng_own_2014_orig_acs
        
        chng_own_tbl_guide <- 
                data.frame(terse = acs.colnames(chng_own_2014_acs_1),
                       pretty = acs.colnames(acs.fetch(endyear = 2014, 
                          geography = geo.make(us = TRUE), 
                          table.number = chng_own_tbl,col.names = 'pretty')))
        # Pull out the specific columns that are important
        
        get_col <- function(pattern){
                chng_own_tbl_guide %>% filter(str_detect(pretty,pattern)) %>% select(terse) %>% unlist(use.names = FALSE) %>% as.character()
        }
        
        total <- get_col('Total')
        
        own <- get_col('Owner\\soccupied:$')
        
        # Find the proportion of Whites to the Total population
        
        acs_col_pct_chng_own_2014 <-  
                acs::apply(X = chng_own_2014_acs_1[,own],
                           MARGIN = 1,
                           FUN = divide.acs,
                           denominator = chng_own_2014_acs_1[,total],
                           method = "proportion",
                           verbose = FALSE)
        acs.colnames(acs_col_pct_chng_own_2014) <- 'PCT_OWN'
        
        # Save the object

acs_col_pct_chng_own_2014 %>% saveRDS(root_file('1-data/4-interim/chng-own-2014-acs.rds'))
}

chng_own_2014_acs <- read_rds(root_file('1-data/4-interim/chng-own-2014-acs.rds'))


# Plot the data
# -----------------------------------------------------------------------------

plot(chng_own_2014_acs[order(estimate(chng_own_2014_acs)),1])


```

```{r chng-own-2009-acs}

# Download the Census table for the project's geographies, 2005-2009
# -----------------------------------------------------------------------------

coo_geo_2009_acs <- read_rds(root_file('1-data/4-interim/coo-geo-2009-acs.rds'))

# 2005-2009, 5-year estimates
if(!file.exists(root_file('1-data/4-interim/chng-own-2009-orig-acs.rds'))){
        
        acs.fetch(endyear = 2009, 
                  geography = coo_geo_2009_acs, 
                  table.number = chng_own_tbl) %>%
                write_rds(root_file('1-data/4-interim/chng-own-2009-orig-acs.rds'))
        
}

chng_own_2009_orig_acs <- read_rds(root_file('1-data/4-interim/chng-own-2009-orig-acs.rds'))

# Download the 2000-2010 Census Tract crosswalk
# and create subsets of the tracts to be transformed
# -----------------------------------------------------------------------------

if(!file.exists(root_file('1-data/4-interim/cw.rds'))){
        read_csv(
                root_file('1-data/3-external/crosswalk_2000_2010.csv'),
                col_types = cols(cbsa10 = col_character(),
                                 ccflag10 = col_character(),
                                 changetype = col_character(),
                                 metdiv10 = col_character(),
                                 placefp10 = col_character())
        ) %>%
                rename(GEOID_2000 = trtid00,
                       GEOID_2010 = trtid10) %>% 
                mutate(KC = str_sub(GEOID_2000,1,5)) %>%
                filter(KC == '53033') %>%
                mutate(GEOID6_2000 = str_sub(GEOID_2000,6,11),
                       GEOID6_2010 = str_sub(GEOID_2010,6,11)) %>% 
                select(matches('GEOID'),everything(),-KC) %>% 
                write_rds(root_file('1-data/4-interim/cw.rds'))
}

cw <- read_rds(root_file('1-data/4-interim/cw.rds'))

# Create a vector of the unchanged tracts
tr_unchng <-
        cw %>%
        filter(changetype %in% '1') %>%
        select(GEOID6_2000) %>%
        unlist(use.names = FALSE)


# Create a vector of the changed tracts
tr_chng_all <-
        cw %>%
        filter(changetype %!in% '1') %>%
        select(GEOID6_2000) %>%
        unlist(use.names = FALSE)


# Create a vector of the consolidated tracts
tr_chng_consl <-
        cw %>%
        filter(changetype %in% '2') %>%
        select(GEOID6_2000) %>%
        unlist(use.names = FALSE)

# Create a vector of the split
tr_chng_split_2000 <- 
        cw %>%
        filter(changetype %in% c('3')) %>%
        select(GEOID6_2000) %>%
        unlist(use.names = FALSE)

tr_chng_split_2010 <- 
        cw %>%
        filter(changetype %in% c('3')) %>%
        select(GEOID6_2010) %>%
        unlist(use.names = FALSE)

# Create a vector of the many-to-many
tr_chng_m2m_2000 <- 
        cw %>%
        filter(changetype %in% c('4')) %>%
        select(GEOID6_2000) %>%
        unlist(use.names = FALSE)

tr_chng_m2m_2010 <- 
        cw %>%
        filter(changetype %in% c('4')) %>%
        select(GEOID6_2010) %>%
        unlist(use.names = FALSE)


# Unchanged tracts
# -----------------------------------------------------------------------------

chng_own_2009_unchng_acs <- 
        chng_own_2009_orig_acs[geography(chng_own_2009_orig_acs)[['tract']] %in% tr_unchng,]


# Changed tracts
# -----------------------------------------------------------------------------

# Consolidated

chng_own_2009_consl_acs <- 
        chng_own_2009_orig_acs[geography(chng_own_2009_orig_acs)[['tract']] %in% tr_chng_consl,]


# Split 

chng_own_2009_split_acs <- 
        chng_own_2009_orig_acs[geography(chng_own_2009_orig_acs)[['tract']] %in% tr_chng_split_2000,]

# Many-to-many

chng_own_2009_m2m_acs <- 
        chng_own_2009_orig_acs[geography(chng_own_2009_orig_acs)[['tract']] %in% tr_chng_m2m_2000,]


# Transformation of changed tracts
# -----------------------------------------------------------------------------

estimate_consl <- function(acs_consl){
        
        if(!exists('cw')){cw <- read_rds(root_file('1-data/4-interim/cw.rds'))}
        
        if(!exists('chng_own_2014_orig_acs')){chng_own_2014_orig_acs <- read_rds(root_file('1-data/4-interim/chng-own-2014-orig-acs.rds'))}
        
        acs_consl_transformed <- acs::apply(X = acs_consl,MARGIN = 1,FUN = sum, verbose = TRUE)
        
        tracts <- chng_own_2009_consl_acs %>% geography() %>% select(tract) %>% unlist(use.names = FALSE)
        
        tr_2010 <- 
                cw %>% 
                select(matches('GEOID|changetype')) %>% 
                filter(GEOID6_2000 %in% tracts) %>% 
                select(GEOID6_2010) %>% 
                unlist(use.names = FALSE) %>% 
                unique
        
        acs_geo_2010 <- 
                chng_own_2014_orig_acs %>% 
                geography() %>% 
                filter(tract %in% tr_2010)
        
        geography(acs_consl_transformed) <- acs_geo_2010
        
        return(acs_consl_transformed)
      
}

# Consolidated

chng_own_2009_consl_transformed_acs <- 
        chng_own_2009_consl_acs %>% 
        estimate_consl()

# Split 

estimate_split <- function(acs_GEOID6_2010){
        
        if(!exists('cw')){cw <- read_rds(root_file('1-data/4-interim/cw.rds'))}
        
        if(!exists('chng_own_2014_orig_acs')){chng_own_2014_orig_acs <- read_rds(root_file('1-data/4-interim/chng-own-2014-orig-acs.rds'))}
        
        acs_weight <- 
                cw %>% 
                filter(GEOID6_2010 %in% acs_GEOID6_2010) %>% 
                select(weight) %>% 
                unlist(use.names = FALSE)
        
        acs_GEOID6_2000 <- 
                cw %>% 
                filter(GEOID6_2010 %in% acs_GEOID6_2010) %>% 
                select(GEOID6_2000) %>% 
                unlist(use.names = FALSE)
        
        acs_split_transformed <- chng_own_2009_split_acs[geography(chng_own_2009_split_acs)$tract %in% acs_GEOID6_2000,] * acs_weight
        
        acs_geo_2010 <- 
                chng_own_2014_acs %>% 
                geography() %>% 
                filter(tract %in% acs_GEOID6_2010)
        
        geography(acs_split_transformed) <- acs_geo_2010
        
        return(acs_split_transformed)
        
}

chng_own_2009_split_transformed_acs <-
        lapply(as.list(tr_chng_split_2010),estimate_split) %>% 
        Reduce('rbind.acs',.)

# Many-to-many
        
estimate_m2m <- function(acs_GEOID6_2010){
        
        if(!exists('cw')){cw <- read_rds(root_file('1-data/4-interim/cw.rds'))}
        
        if(!exists('chng_own_2014_orig_acs')){chng_own_2014_orig_acs <- read_rds(root_file('1-data/4-interim/chng-own-2014-orig-acs.rds'))}
        
        acs_GEOID6_2000 <- 
                cw %>% 
                filter(GEOID6_2010 %in% acs_GEOID6_2010) %>% 
                arrange(GEOID_2010,GEOID_2000) %>% 
                select(GEOID6_2000) %>% 
                unlist(use.names = FALSE)
        
        acs_weight_2000 <- 
                cw %>% 
                filter(GEOID6_2010 %in% acs_GEOID6_2010) %>% 
                arrange(GEOID_2010,GEOID_2000) %>% 
                select(weight) %>% 
                unlist(use.names = FALSE)
        
        tmp <- 
                chng_own_2009_m2m_acs[geography(chng_own_2009_m2m_acs)$tract %in% acs_GEOID6_2000,]

        acs_m2m_transformed <- 
                acs::apply(tmp,MARGIN = 1,FUN = function(x) x*acs_weight_2000) %>%
                acs::apply(MARGIN = 1,FUN = sum)
        
        acs_geo_2010 <- 
                chng_own_2014_acs %>% 
                geography() %>% 
                filter(tract %in% acs_GEOID6_2010)
        
        geography(acs_m2m_transformed) <- acs_geo_2010
        
        return(acs_m2m_transformed)
        
}

chng_own_2009_m2m_transformed_acs <- 
        lapply(as.list(unique(tr_chng_m2m_2010)), estimate_m2m) %>% 
        Reduce('rbind.acs',.)

# Combine KC, Seattle CCD, and all 2004-2009 tracts (unchanged and transformed)
# -----------------------------------------------------------------------------

chng_own_2009_tr_transformed_acs <- 
        Reduce('rbind.acs',list(
                chng_own_2009_unchng_acs,
                chng_own_2009_consl_transformed_acs,
                chng_own_2009_split_transformed_acs,
                chng_own_2009_m2m_transformed_acs
        )) %>% 
        .[order(geography(.)$tract),] %>%  # order by tract
        .[!duplicated(geography(.)$tract),] # remove duplicates


# Combine tracts to create the COO Communities
# -----------------------------------------------------------------------------

coo_geos_sf <- read_rds(root_file('1-data/4-interim/coo-geos-sf.rds')) 

make_comm <- function(tr_sf,nm_comm,nm_geog,acs_2009,acs_2014){
        tr <- tr_sf %>% filter(NAME %in% nm_comm & !is.na(GEOID6)) %>% select(GEOID6) %>% unlist(use.names = FALSE)
        
        comm <- acs::apply(X = acs_2009[geography(acs_2009)$tract %in% tr],MARGIN = 1,FUN = sum)
        
        geography(comm) <- geography(acs_2014)[geography(acs_2014)$NAME %in% nm_geog,]
                
        return(comm)
   
}

comm_rv_2009 <- make_comm(tr_sf = coo_geos_sf,
                 nm_comm = 'RV',
                 nm_geog = 'RV',
                 acs_2009 = chng_own_2009_tr_transformed_acs,
                 acs_2014 = chng_own_2014_acs)

comm_wc_2009 <- make_comm(tr_sf = coo_geos_sf,
                 nm_comm = 'WC',
                 nm_geog = 'WC',
                 acs_2009 = chng_own_2009_tr_transformed_acs,
                 acs_2014 = chng_own_2014_acs)

comm_stkw_2009 <- make_comm(tr_sf = coo_geos_sf,
                 nm_comm = 'STKW',
                 nm_geog = 'STKW',
                 acs_2009 = chng_own_2009_tr_transformed_acs,
                 acs_2014 = chng_own_2014_acs)

comm_coo_2009 <- make_comm(tr_sf = coo_geos_sf,
                 nm_comm = c('RV','WC','STKW'),
                 nm_geog = 'COO',
                 acs_2009 = chng_own_2009_tr_transformed_acs,
                 acs_2014 = chng_own_2014_acs)


# Combine all 2004-2009 geographies
# -----------------------------------------------------------------------------


chng_own_2009_all_geog <- 
        Reduce('rbind.acs',list(
                chng_own_2009_tr_transformed_acs,
                comm_rv_2009,
                comm_wc_2009,
                comm_stkw_2009,
                comm_coo_2009,
                chng_own_2009_orig_acs[2:1,]
        )) 



# Process and save the data
# -----------------------------------------------------------------------------

if(!file.exists(root_file('1-data/4-interim/chng-own-2009-acs.rds'))){
        chng_own_2009_acs_1 <- chng_own_2009_all_geog
        
        chng_own_tbl_guide <- 
                data.frame(terse = acs.colnames(chng_own_2009_acs_1),
                       pretty = acs.colnames(acs.fetch(endyear = 2009, 
                          geography = geo.make(us = TRUE), 
                          table.number = chng_own_tbl,col.names = 'pretty')))
        # Pull out the specific columns that are important
        
        get_col <- function(pattern){
                chng_own_tbl_guide %>% filter(str_detect(pretty,pattern)) %>% select(terse) %>% unlist(use.names = FALSE) %>% as.character()
        }
        
        total <- get_col('Total')
        
        own <- get_col('Owner\\soccupied:$')
        
        # Find the proportion of Whites to the Total population
        
        acs_col_pct_chng_own_2009 <-  
                acs::apply(X = chng_own_2009_acs_1[,own],
                           MARGIN = 1,
                           FUN = divide.acs,
                           denominator = chng_own_2009_acs_1[,total],
                           method = "proportion",
                           verbose = FALSE)
        
        acs.colnames(acs_col_pct_chng_own_2009) <- 'PCT_OWN'
        
        # Save the object

acs_col_pct_chng_own_2009 %>% write_rds(root_file('1-data/4-interim/chng-own-2009-acs.rds'))
}

chng_own_2009_acs <- read_rds(root_file('1-data/4-interim/chng-own-2009-acs.rds'))


# Plot the data
# -----------------------------------------------------------------------------

plot(chng_own_2009_acs[order(estimate(chng_own_2009_acs)),1])



```

```{r chng-own-acs}

if(!file.exists(root_file('1-data/4-interim/chng-own-acs.rds'))){
        # remove any unshared geographies

in_2009_not_2014 <- 
        anti_join(geography(chng_own_2009_acs),
                              geography(chng_own_2014_acs),
                              by = 'NAME') %>% 
        select(tract) %>% 
        unlist(use.names = FALSE)

in_2014_not_2009 <- anti_join(geography(chng_own_2014_acs),
                              geography(chng_own_2009_acs),
                              by = 'NAME') %>% 
        select(tract) %>% 
        unlist(use.names = FALSE)


chng_own_2014_final_acs <- chng_own_2014_acs[geography(chng_own_2014_acs)$tract %!in% in_2014_not_2009,] %>% 
        .[order(geography(.)$tract),]

chng_own_2009_final_acs <- chng_own_2009_acs[geography(chng_own_2009_acs)$tract %!in% in_2009_not_2014,] %>% 
        .[order(geography(.)$tract),]

chng_own_acs <- chng_own_2014_final_acs - chng_own_2009_final_acs

acs.colnames(chng_own_acs) <- 'CHNG_PCT_OWN'

check_names <- chng_own_acs %>% geography() %>% .$NAME

geography(chng_own_acs) <- geography(chng_own_2014_final_acs)

chng_own_acs %>% write_rds(root_file('1-data/4-interim/chng-own-acs.rds'))

}

chng_own_acs <-  read_rds(root_file('1-data/4-interim/chng-own-acs.rds'))


```



```{r chng-own-sf}

if(!file.exists(root_file('1-data/4-interim/chng-own-sf.rds'))){
        # Read in the `sf` object and re-order the records to by `JOIN_ID` 
# -----------------------------------------------------------------------------

coo_geos_sf <- 
        read_rds(root_file('1-data/4-interim/coo-geos-sf.rds'))

coo_geos_sf_ordered <- coo_geos_sf %>% arrange(desc(JOIN_ID)) %>% select(-JOIN_ID)

#
# Convert the `acs` object into a dataframe, join to sf object
# -----------------------------------------------------------------------------

coo_chng_own_sf <- 
        data.frame(
                geography(chng_own_acs)["NAME"],
                geography(chng_own_acs)["tract"],
                estimate(chng_own_acs), 
                1.645 * standard.error(chng_own_acs)) %>% 
        `colnames<-`(., c("NAME","GEOID6", "CHNG_PCT_OWN_EST","CHNG_PCT_OWN_MOE")) %>% 
        mutate(JOIN_ID = case_when(.$NAME %in% 'King County, Washington' ~ 'KC',
                                  .$NAME %in% 'Seattle CCD, King County, Washington' ~ 'SEACCD',
                                  !is.na(.$GEOID6) ~ .$GEOID6),
               UPPER = CHNG_PCT_OWN_EST + CHNG_PCT_OWN_MOE, 
               LOWER = CHNG_PCT_OWN_EST - CHNG_PCT_OWN_MOE, 
               POC_PCT_UPPER = if_else(UPPER > 1, 1, UPPER), 
               POC_PCT_LOWER = if_else(LOWER < 0, 0, LOWER)) %>% 
        select(JOIN_ID,everything(),-NAME,-GEOID6,-UPPER,-LOWER) %>% 
        arrange(desc(JOIN_ID)) %>%
        bind_cols(coo_geos_sf_ordered) %>% 
        st_sf()

#
# Save the object
# -----------------------------------------------------------------------------
coo_chng_own_sf %>% write_rds(root_file('1-data/4-interim/chng-own-sf.rds'))
}

coo_chng_own_sf <-  read_rds(root_file('1-data/4-interim/chng-own-sf.rds'))


```

#### Maps

_Add a description sentence_

##### Seattle CCD 
```{r chng-own-seaccd-map}

tr <- 
        coo_chng_own_sf %>% 
        filter(GEOGRAPHY == 'tract' & SEACCD_LGL)

tr_comm <- 
        coo_chng_own_sf %>% 
        filter(GEOGRAPHY == 'tract' & SEACCD_LGL &!is.na(NAME))

comm <- coo_chng_own_sf %>% 
        filter(GEOGRAPHY == 'community')

comm_cntr <- comm %>% mutate(geometry = st_centroid(comm))

pal_chng_own <- colorNumeric(viridis(256, option = "C"),tr$CHNG_PCT_OWN_EST)

myLfltGrey() %>% 
        addPolygons(data = as(tr,'Spatial'),
                fillOpacity = .75,
                fillColor = ~pal_chng_own(CHNG_PCT_OWN_EST),
                color = col2hex('white'),
                opacity = 1,
                weight = .5,
                smoothFactor = 0) %>%
        addPolygons(data = as(comm,'Spatial'),
                fillOpacity = 0,
                color = proj_dark_grey,
                opacity = 1,
                weight = 2,
                smoothFactor = 0) %>%
        addAwesomeMarkers(data = as(comm_cntr, "Spatial"),options = markerOptions(opacity = 0),label = ~ NAME_FULL, labelOptions = labelOptions(noHide = TRUE,direction = 'left', textOnly = TRUE)) %>% 
        addLegend(title = "Change in<br>% Home Ownership",
                  position = 'topright',
                  pal = pal_chng_own, 
                  values = tr$CHNG_PCT_OWN_EST,
                  opacity = .75,
                  labFormat = labelFormat(suffix = '%')) %>% 
        styleWidget(style = "text-transform:uppercase;text-shadow:-1px 0 #FFFFFF,0 1px #FFFFFF,1px 0 #FFFFFF,0 -1px #FFFFFF")

```

#####COO Communities
```{r chng-own-comm-map}
myLfltGrey() %>% 
        addPolygons(data = as(tr_comm,'Spatial'),
                fillOpacity = .75,
                fillColor = ~pal_chng_own(CHNG_PCT_OWN_EST),
                color = col2hex('white'),
                opacity = 1,
                weight = .5,
                smoothFactor = 0) %>%
        addPolygons(data = as(comm,'Spatial'),
                fillOpacity = 0,
                color = proj_dark_grey,
                opacity = 1,
                weight = 2,
                smoothFactor = 0) %>%
        addAwesomeMarkers(data = as(comm_cntr, "Spatial"),options = markerOptions(opacity = 0),label = ~ NAME_FULL, labelOptions = labelOptions(noHide = TRUE,direction = 'left', textOnly = TRUE)) %>% 
        addLegend(title = "Change in<br>% Home Ownership",
                  position = 'topright',
                  pal = pal_chng_own, 
                  values = tr$CHNG_PCT_OWN_EST,
                  opacity = .75,
                  labFormat = labelFormat(suffix = '%')) %>% 
        styleWidget(style = "text-transform:uppercase;text-shadow:-1px 0 #FFFFFF,0 1px #FFFFFF,1px 0 #FFFFFF,0 -1px #FFFFFF")

```
