---
df_print: tibble
output:
  html_notebook:
    code_folding: hide
  pdf_document:
    keep_tex: yes
always_allow_html: yes
---

```{r vuln-low-inc-setup, echo = FALSE, warning=FALSE,message=FALSE,comment=FALSE}
library(plyr)
library(knitr)
library(rprojroot)
library(rgdal)
library(sp)
library(rgeos)
library(tigris)
library(leaflet)
library(ggthemes)
library(magrittr)
library(stringr)
library(downloader)
library(webshot)
library(htmltools)
library(gplots)
library(ggmap)
library(shiny)
library(htmlwidgets)
library(readxl)
library(acs)
library(RColorBrewer)
library(tidyverse)
library(miscgis)
library(operator.tools)
library(leaflet.extras)
library(viridisLite)
library(sf)
root <- rprojroot::is_rstudio_project
root_file <- root$make_fix_file()
opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, comment=FALSE)
```

```{r vuln-low-inc-colors}
green <- miscgis::miscgis_pals$tableau_cat[["green"]]
blue <- miscgis::miscgis_pals$tableau_cat[["blue"]]
orange <- miscgis::miscgis_pals$tableau_cat[["orange"]]
red <- miscgis::miscgis_pals$tableau_cat[["red"]]
teal <- miscgis::miscgis_pals$tableau_cat[["teal"]]
pal_rgb_4 <- miscgis::miscgis_pals$tableau_cat[c("red","gold","green","blue")] %>% unlist %>% palette()
pal_rgb_4 <- miscgis::miscgis_pals$tableau_cat[c("red","gold","green","blue")] %>% unlist %>% palette()
pal_rgb_6 <- miscgis::miscgis_pals$tableau_cat[c("red","gold","green","blue","orange","purple")] %>% unlist %>% palette()
pal_rgb_6 <- miscgis::miscgis_pals$tableau_cat[c("red","gold","green","blue","orange","purple")] %>% unlist %>% palette()
proj_light_grey <- col2hex("grey75")
proj_grey <- col2hex("grey50")
proj_dark_grey <- col2hex("grey25")
```

### Income: Low Income {-}

This indicator is intended as a proxy for low-income households. There are many possible ways to determine whether a household fits this category, but one commonly-used metric is the area median income (AMI). Generally speaking, households earning less than 80% AMI are considered to be low-income households, as status which has implications for affordable housing eligibility. 

The Lisa Bates used a fairly precise dataset for this indicator: percent of households that have incomes at or below at or below 80% of the HUD-adjusted median family income (MFI). The data source is HUD's Consolidated Housing Affordability Strategy (CHAS) as used. Unfortunately, the [most recent CHAS data](https://www.huduser.gov/portal/datasets/cp/CHAS/data_download_chas.html) is based on the 2009-2013 ACS 5-year estimate. 

Rather than introduce an inconsistency in the observation periods of this assessment's datasets, this method uses an approximation of the 80% AMI metric using the ACS Household Income (B19001), 2010-2014 5-year estimate. This data is structured as household counts within a range of annual household income (e.g. $15,000 to $19,999). The FY 2015 MFI Estimate for Seattle-Bellevue, WA HUD Metro FMR Area is $89,500. In inflation-adjusted 2014 dollars that is $87,732.78, and 80% of that equals $70,200; which puts it within the `$60,000 to $74,999` range. Therefore, for this assessment any household with an annual income of $74,999 or less is considered to be low-income.

```{r vuln-low-inc-acs, fig.cap= 'Households earning less than 80% of the HUD, King County Census Tracts, 2010-2014'}

# Download the Census table for the project's geographies
# -----------------------------------------------------------------------------

if(!exists('coo_geo_acs')){coo_geo_acs <- read_rds(root_file('1-data/4-interim/coo-geo-acs.rds'))}

vuln_low_inc_tbl <- "B19001"  # census table code

if(!file.exists(root_file('1-data/4-interim/vuln-low-inc-orig-acs.rds'))){
        
        acs.fetch(endyear = 2014, geography = coo_geo_acs, 
                  table.number = vuln_low_inc_tbl) %>%
                write_rds(root_file('1-data/4-interim/vuln-low-inc-orig-acs.rds'))
        
}

vuln_low_inc_orig_acs <- read_rds(root_file('1-data/4-interim/vuln-low-inc-orig-acs.rds'))


# Process and save the data
# -----------------------------------------------------------------------------

if(!file.exists(root_file('1-data/4-interim/vuln-low-inc-acs.rds'))){
        vuln_low_inc_acs1 <- vuln_low_inc_orig_acs
        
        vuln_low_inc_tbl_guide <- 
                data.frame(terse = acs.colnames(vuln_low_inc_acs1),
                       pretty = acs.colnames(acs.fetch(endyear = 2014, 
                          geography = geo.make(us = TRUE), 
                          table.number = vuln_low_inc_tbl,col.names = 'pretty')))
        # Pull out the specific columns that are important
        
        get_col <- function(pattern){
                vuln_low_inc_tbl_guide %>% filter(str_detect(pretty,pattern)) %>% select(terse) %>% unlist(use.names = FALSE) %>% as.character()
        }
        
        # Indicator universe total
        
        total <- get_col('Total:')
        
        vuln_low_total_inc <- vuln_low_inc_acs1[,total]
        
        acs.colnames(vuln_low_total_inc) <- "HH_TOTAL_CNT"
        
        
        # Indicator count (Add all pops with bach. or higher)
        
        low_inc <- get_col('\\$10,|\\$14,|\\$19,|\\$24,|\\$29,|\\$34,|\\$39,|\\$44,|\\$49,|\\$59,|\\$74,')
        
        acs_col_low_inc <- 
                acs::apply(X = vuln_low_inc_acs1[,low_inc],
                           MARGIN = 2,
                           FUN = sum,
                           verbose = FALSE)
        
        acs.colnames(acs_col_low_inc) <- "LOW_INC_CNT"
        
        
        # Indicator percent (the proportion of low income households)
        
        acs_col_pct_low_inc <-  
                acs::apply(X = acs_col_low_inc,
                           MARGIN = 1,
                           FUN = divide.acs,
                           denominator = vuln_low_total_inc,
                           method = "proportion",
                           verbose = FALSE)
        
        acs.colnames(acs_col_pct_low_inc) <- 'LOW_INC_PCT'
        
        vuln_low_inc_acs <- Reduce('cbind',c(acs_col_pct_low_inc,acs_col_low_inc,vuln_low_total_inc))
        
        
        # Save the object

vuln_low_inc_acs %>% saveRDS(root_file('1-data/4-interim/vuln-low-inc-acs.rds'))

}

vuln_low_inc_acs <- read_rds(root_file('1-data/4-interim/vuln-low-inc-acs.rds'))


# Plot the data
# -----------------------------------------------------------------------------

# Sanity check: the University of Washington Tract appears to have a very high margin of error (likely due to the small sample size), so it should be removed

vuln_low_inc_noUW_acs <- vuln_low_inc_acs[standard.error(vuln_low_inc_acs[,'LOW_INC_PCT'])<.15,'LOW_INC_PCT']

# For some reason the column name gets dropped, and it needs to be redefined before 
# the object is passed to plot() 

acs.colnames(vuln_low_inc_noUW_acs) <- 'LOW_INC_PCT'

par(mar = c(1,4,1,1))
plot(vuln_low_inc_noUW_acs[order(estimate(vuln_low_inc_noUW_acs)),1],ylab = 'People of Color (%)',yaxt = 'n',xaxt = 'n',bty = 'n')
yticks_val <- scales::pretty_breaks(n=5)(vuln_low_inc_noUW_acs@estimate)
yticks <- scales::pretty_breaks(n=5)(vuln_low_inc_noUW_acs@estimate)
axis(2, at=yticks_val,labels = 100*yticks)


```

```{r vuln-low-inc-sf}

if(!file.exists(root_file('1-data/4-interim/vuln-low-inc-sf.rds'))){
        # Read in the `sf` object and re-order the records to by `JOIN_ID` 
# -----------------------------------------------------------------------------

if(!exists('coo_geos_sf')){coo_geos_sf <- read_rds(root_file('1-data/4-interim/coo-geos-sf.rds'))}    

#
# Convert the `acs` object into a dataframe, join to sf object
# -----------------------------------------------------------------------------
coo_vuln_low_inc_sf <- 
        data.frame(
                geography(vuln_low_inc_acs)["NAME"],
                geography(vuln_low_inc_acs)["tract"],
                estimate(vuln_low_inc_acs), 
                1.645 * standard.error(vuln_low_inc_acs)) %>% 
        `colnames<-`(., c("NAME","GEOID6","LOW_INC_PCT_EST","LOW_INC_CNT_EST","HH_TOTAL_CNT_EST","LOW_INC_PCT_MOE","LOW_INC_CNT_MOE","HH_TOTAL_CNT_MOE")) %>% 
        mutate(JOIN_ID = case_when(.$NAME %in% 'King County, Washington' ~ 'KC',
                                   .$NAME %in% 'Seattle CCD, King County, Washington' ~ 'SEACCD',
                                   !is.na(.$GEOID6) ~ .$GEOID6)) %>%
        gather(VAR_TYPE,VALUE, matches('EST|MOE')) %>% 
        mutate(VAR = str_replace_all(VAR_TYPE,'_EST|_MOE',''),
               TYPE = str_extract_all(VAR_TYPE,'EST|MOE',simplify = FALSE) %>% unlist(use.names = FALSE)) %>%
        spread(TYPE,VALUE) %>%
        arrange(NAME,VAR) %>%
        group_by(NAME,VAR) %>% 
        select(-VAR_TYPE) %>% 
        summarise_all(first_not_na) %>% 
        mutate(UPPER = EST + MOE,
               LOWER = EST - MOE) %>%
        ungroup %>% 
        gather(TYPE,VAL,EST:LOWER) %>% 
        mutate(VAR_TYPE = paste(VAR,TYPE,sep = "_")) %>% 
        arrange(VAR_TYPE) %>% 
        spread(VAR_TYPE,VAL) %>% 
        group_by(NAME,VAR) %>% 
        summarise_all(first_not_na) %>% 
        ungroup %>% 
        select(-VAR,-TYPE) %>% 
        group_by(NAME) %>% 
        summarise_all(first_not_na) %>% 
        select(JOIN_ID,
               matches('LOW_INC_PCT'),
               matches('LOW_INC_CNT'),
               everything(),
               -NAME,-GEOID6) %>% as.data.frame() %>% 
        full_join(coo_geos_sf,by = 'JOIN_ID') %>% 
        st_sf()
# coo_vuln_low_inc_sf <- 
#         data.frame(
#                 geography(vuln_low_inc_acs)["NAME"],
#                 geography(vuln_low_inc_acs)["tract"],
#                 estimate(vuln_low_inc_acs), 
#                 1.645 * standard.error(vuln_low_inc_acs)) %>% 
#         `colnames<-`(., c("NAME","GEOID6", "LOW_INC_PCT_EST","LOW_INC_PCT_MOE")) %>% 
#         mutate(JOIN_ID = case_when(.$NAME %in% 'King County, Washington' ~ 'KC',
#                                   .$NAME %in% 'Seattle CCD, King County, Washington' ~ 'SEACCD',
#                                   !is.na(.$GEOID6) ~ .$GEOID6),
#                UPPER = LOW_INC_PCT_EST + LOW_INC_PCT_MOE, 
#                LOWER = LOW_INC_PCT_EST - LOW_INC_PCT_MOE, 
#                LOW_INC_PCT_UPPER = if_else(UPPER > 1, 1, UPPER), 
#                LOW_INC_PCT_LOWER = if_else(LOWER < 0, 0, LOWER)) %>% 
#         select(JOIN_ID,everything(),-NAME,-GEOID6,-UPPER,-LOWER) %>% 
#         arrange(desc(JOIN_ID)) %>%
#         bind_cols(coo_geos_sf_ordered) %>% 
#         st_sf()

#
# Save the object
# -----------------------------------------------------------------------------
coo_vuln_low_inc_sf %>% write_rds(root_file('1-data/4-interim/vuln-low-inc-sf.rds'))
}

coo_vuln_low_inc_sf <-  read_rds(root_file('1-data/4-interim/vuln-low-inc-sf.rds'))

```

##### Seattle CCD {-}
```{r vuln-low-inc-seaccd-map}

tr <- 
        coo_vuln_low_inc_sf %>% 
        filter(GEOGRAPHY == 'tract' & SEACCD_LGL)

seaccd_row <- which(coo_vuln_low_inc_sf$JOIN_ID %in% 'SEACCD')[[1]]

my_pct <- function(x) round_any(x*100,1) %>% paste0("%")

popup_tbl_df <- 
        coo_vuln_low_inc_sf %>%
        mutate(LOW_INC_PCT_EST = my_pct(LOW_INC_PCT_EST),
               LOW_INC_CNT_EST = scales::comma(LOW_INC_CNT_EST),
               HH_TOTAL_CNT_EST = scales::comma(HH_TOTAL_CNT_EST)) %>% 
        select(-geometry) %>% 
        select('%' = LOW_INC_PCT_EST,
               '#' = LOW_INC_CNT_EST,
               '#' = HH_TOTAL_CNT_EST) %>% 
        as.data.frame(check.names = FALSE) %>% 
        `row.names<-`(coo_vuln_low_inc_sf$JOIN_ID)

if(!file.exists(root_file('1-data/4-interim/vuln-low-inc-seaccd-popup-tbl.rds'))){
        lapply(seq_along(1:nrow(popup_tbl_df)), 
               function(row) htmlTable::htmlTable(popup_tbl_df[c(row,seaccd_row),],
                                                  cgroup = c("Low Inc", "Total HH"),n.cgroup = c(2,1)))[coo_vuln_low_inc_sf$GEOGRAPHY %in% 'tract' & coo_vuln_low_inc_sf$SEACCD_LGL] %>% 
                write_rds(root_file('1-data/4-interim/vuln-low-inc-seaccd-popup-tbl.rds'))
}

if(!file.exists(root_file('1-data/4-interim/vuln-low-inc-comm-popup-tbl.rds'))){
        lapply(seq_along(1:nrow(popup_tbl_df)), 
               function(row) htmlTable::htmlTable(popup_tbl_df[c(row,seaccd_row),],
                                                  cgroup = c("Low Inc", "Total HH"),n.cgroup = c(2,1)))[coo_vuln_low_inc_sf$GEOGRAPHY %in% 'tract' & coo_vuln_low_inc_sf$SEACCD_LGL &!is.na(coo_vuln_low_inc_sf$NAME)] %>% 
                write_rds(root_file('1-data/4-interim/vuln-low-inc-comm-popup-tbl.rds'))
}

popup_tbl_seaccd <- read_rds(root_file('1-data/4-interim/vuln-low-inc-seaccd-popup-tbl.rds'))

tr_comm <- 
        coo_vuln_low_inc_sf %>% 
        filter(GEOGRAPHY == 'tract' & SEACCD_LGL &!is.na(NAME))

comm <- coo_vuln_low_inc_sf %>% 
        filter(GEOGRAPHY == 'community')

comm_cntr <- comm %>% mutate(geometry = st_centroid(comm))

comm_bounds <- comm %>% 
        as("Spatial") %>% 
        as("SpatialLines")

pal_low_inc <- colorNumeric(viridis(256, option = "C"),tr$LOW_INC_PCT_EST)

myLfltGrey() %>% 
        myLfltOpts() %>% 
        addPolygons(data = as(tr,'Spatial'),
                fillOpacity = .75,
                fillColor = ~pal_low_inc(LOW_INC_PCT_EST),
                color = col2hex('white'),
                opacity = 1,
                weight = .5,
                smoothFactor = 0,
                popup = popup_tbl_seaccd) %>%
        addPolylines(data = comm_bounds,
                fillOpacity = 0,
                color = proj_dark_grey,
                opacity = 1,
                weight = 2,
                smoothFactor = 0) %>%
        addLabelOnlyMarkers(data = as(comm_cntr, "Spatial"),label = ~ NAME_FULL, labelOptions = labelOptions(noHide = TRUE,direction = 'left', textOnly = TRUE)) %>% 
        addLegend(title = "Low Income<br>Households",
                  position = 'topright',
                  pal = pal_low_inc, 
                  values = tr$LOW_INC_PCT_EST,
                  opacity = .75,
                  labFormat = labelFormat(suffix = '%',transform = function(x) 100 * x)) %>% 
        styleWidget(style = "text-transform:uppercase;text-shadow:-1px 0 #FFFFFF,0 1px #FFFFFF,1px 0 #FFFFFF,0 -1px #FFFFFF")

```

#####COO Communities {-}
```{r vuln-low-inc-comm-map}

popup_tbl_comm <- read_rds(root_file('1-data/4-interim/vuln-low-inc-comm-popup-tbl.rds'))

myLfltGrey() %>% 
        myLfltOpts() %>% 
        addPolygons(data = as(tr_comm,'Spatial'),
                fillOpacity = .75,
                fillColor = ~pal_low_inc(LOW_INC_PCT_EST),
                color = col2hex('white'),
                opacity = 1,
                weight = .5,
                smoothFactor = 0,
                popup = popup_tbl_comm) %>%
        addPolylines(data = comm_bounds,
                fillOpacity = 0,
                color = proj_dark_grey,
                opacity = 1,
                weight = 2,
                smoothFactor = 0) %>%
        addLabelOnlyMarkers(data = as(comm_cntr, "Spatial"),label = ~ NAME_FULL, labelOptions = labelOptions(noHide = TRUE,direction = 'left', textOnly = TRUE)) %>% 
        addLegend(title = "Renters",
                  position = 'topright',
                  pal = pal_low_inc, 
                  values = tr$LOW_INC_PCT_EST,
                  opacity = .75,
                  labFormat = labelFormat(suffix = '%',transform = function(x) 100 * x)) %>% 
        styleWidget(style = "text-transform:uppercase;text-shadow:-1px 0 #FFFFFF,0 1px #FFFFFF,1px 0 #FFFFFF,0 -1px #FFFFFF")

```

