---
df_print: tibble
output:
  html_notebook:
    code_folding: hide
  pdf_document:
    keep_tex: yes
always_allow_html: yes
---

```{r hra-tr-setup, echo = FALSE, warning=FALSE,message=FALSE,comment=FALSE}
library(plyr)
library(knitr)
library(rprojroot)
library(tidyverse)
library(rgdal)
library(sp)
library(rgeos)
library(miscgis)
library(tigris)
library(leaflet)
library(ggthemes)
library(magrittr)
library(stringr)
library(downloader)
library(webshot)
library(htmltools)
library(gplots)
library(ggmap)
library(shiny)
library(htmlwidgets)
library(readxl)
library(acs)
library(RColorBrewer)
root <- rprojroot::is_rstudio_project
root_file <- root$make_fix_file()
opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, comment=FALSE)
```

### Introduction
Resolving geographic units that do not neatly coincide is a common problem in spatial data analysis. This method attempts to conflate King County Health Reporting Areas (HRAs) to US Census tracts. In the cases where a given tract with entirely within a HRA, that tract receives the HRA's ID. Where a given tract overlaps multiple HRAs block-level census data is used to determine which HRA ID to assign to the tract. 


### Census Block Counts
This method provides three alternatives of block-level counts that can be used:

  * Population
  * Housing Units
  * Population in Housing Units
  

```{r hra-tr-blks}

if(!file.exists(root_file('1-data/4-interim/kc-blk-sp.gpkg'))){
        pop <- read_csv(root_file('1-data/3-external/manual/wa-blk/DEC_10_SF1_P1/DEC_10_SF1_P1_with_ann.csv'), 
                        col_types = cols(Id2 = col_character()), 
                        skip = 1) %>% 
                mutate(GEO_ID_BLK = Id2,
                       GEOID_TR = substr(Id2,start = 1,stop = 11),
                       POP = Total) %>% 
                select(GEO_ID_BLK,GEOID_TR,POP)
        
        hu <- read_csv(root_file('1-data/3-external/manual/wa-blk/DEC_10_SF1_H1/DEC_10_SF1_H1_with_ann.csv'), 
                       col_types = cols(Id2 = col_character()), 
                       skip = 1) %>% 
                mutate(GEO_ID_BLK = Id2,
                       HU = Total) %>% 
                select(GEO_ID_BLK,HU)
        
        pophu <- read_csv(root_file('1-data/3-external/manual/wa-blk/DEC_10_SF1_H10/DEC_10_SF1_H10_with_ann.csv'), 
                          col_types = cols(Id2 = col_character()), 
                          skip = 1) %>% 
                mutate(GEO_ID_BLK = Id2,
                       POPHU = Total) %>% 
                select(GEO_ID_BLK,POPHU)
        
        cnts <- left_join(pop,hu,by = 'GEO_ID_BLK') %>% 
                left_join(pophu,by = 'GEO_ID_BLK')
        
        # Source for KC Blocks spatial data
        if(!file.exists(root_file('1-data/3-external/wa-blk/blocks10/blocks10.shp'))){
                url <- 'ftp://ftp.kingcounty.gov/gis-web/web/GISData/blocks10_SHP.zip' # direct URL to the file download
                
                temp <- tempfile() # create a temporary file to hold the compressed download
                
                download(url, dest = temp, mode='wb') # download the file
                
                unzip (temp, exdir = root_file('1-data/3-external/wa-blk')) # extract the file to the project folder
        }
        
        kc_blk <- readOGR(dsn = root_file('1-data/3-external/wa-blk/blocks10/'),
                          layer = 'blocks10',
                          verbose = FALSE,
                          stringsAsFactors = FALSE)
        
        kc_blk@data %<>% left_join(cnts, by = 'GEO_ID_BLK') %>% 
                select(GEOID_BLK = GEO_ID_BLK,
                       GEOID_TR:POPHU)
        
        kc_blk %>% writeOGR(dsn = root_file('1-data/4-interim/kc-blk-sp.gpkg'),
                            layer = 'kc_blk_sp',
                            driver = 'GPKG',
                            overwrite_layer = TRUE,verbose = FALSE)
}

kc_blk_sp <- readOGR(dsn = root_file('1-data/4-interim/kc-blk-sp.gpkg'),
                            layer = 'kc_blk_sp',
                  stringsAsFactors = FALSE,
                  verbose = FALSE)



```


### The Algorithm 
The following actions are performed in this method:

  1. Centroids of the census block polygons are calculated (`class = SpatialPointsDataFrame`)
  2. HRA IDs are passed to the block centroid using a spatial overlay method (`sp::over()`)
  3. Blocks are aggregated into tracts and the count variables (`POP`,`HU`,`POPHU`) are summed
  4. For each count variable, the HRA ID with the highest sum is assigned to each tract
  
```{r hra-assign}

# Pass HRA IDs to the block centroids
hra <- readOGR(dsn = root_file('1-data/3-external/manual/HRA_2010Block_Clip/'),layer = 'HRA_2010Block_Clip',
               verbose = FALSE,stringsAsFactors = FALSE)

kc_blk_cnt_sp <- SpatialPointsDataFrame(coords = rgeos::gCentroid(kc_blk_sp,byid = TRUE),
                               data = as.data.frame(kc_blk_sp@data),
                               match.ID = FALSE)

kc_blk_cnt_sp$HRA_ID <- sp::over(kc_blk_cnt_sp,hra[,'HRA2010v2_']) %>% unlist

kc_blk_cnt_sp@data %<>% mutate(HRA_ID = ifelse(is.na(HRA_ID),'None',HRA_ID))


# Assign HRAs to tracts
 
if(!file.exists(root_file('1-data/4-interim/kc-hra-tr-sp.gpkg'))){
        if(!file.exists(root_file('1-data/3-external/kc-tr/tracts10_shore/tracts10_shore.shp'))){
                
                url <- 'ftp://ftp.kingcounty.gov/gis-web/web/GISData/tracts10_shore_SHP.zip' # direct URL to the file download
                
                temp <- tempfile() # create a temporary file to hold the compressed download
                
                download(url, dest = temp, mode='wb') # download the file
                
                unzip (temp, exdir = root_file('1-data/3-external/kc-tr')) # extract the file to the project folder
                
                
        }
         
        kc_tr_sp <- readOGR(dsn = root_file('1-data/3-external/kc-tr/tracts10_shore/'),
                            layer = 'tracts10_shore',
                            verbose = FALSE,
                            stringsAsFactors = FALSE)
        kc_tr_sp@data %<>% select(GEOID_TR = GEO_ID_TRT)
        
        first_notNA <- function(x){first(x[!is.na(x)])}
        
        tr_hra_ids <- 
                kc_blk_cnt_sp@data %>% 
                as.data.frame() %>% 
                gather('VAR','COUNT',POP:POPHU) %>% 
                group_by(GEOID_TR,VAR,HRA_ID) %>% 
                summarise(SUM = sum(COUNT)) %>% 
                arrange(desc(SUM)) %>% 
                slice(1) %>% 
                spread(VAR,HRA_ID) %>% 
                mutate(HU_CNT = ifelse(!is.na(HU),SUM,NA_integer_),
                       POP_CNT = ifelse(!is.na(POP),SUM,NA_integer_),
                       POPHU_CNT = ifelse(!is.na(POPHU),SUM,NA_integer_)) %>% 
                select(-SUM) %>% 
                group_by(GEOID_TR) %>% 
                summarise_all(funs(first_notNA)) %>% 
                ungroup() %>% 
                select(GEOID_TR,
                       POP = POP_CNT,
                       HU = HU_CNT,
                       POPHU = POPHU_CNT,
                       HRA_POP = POP,
                       HRA_HU = HU,
                       HRA_POPHU = POPHU)
        
        tr_hra_ids2 <- 
                tr_hra_ids %>% 
                select(GEOID_TR,matches('HRA')) %>% 
                gather("VAR","HRA",matches('HRA')) %>%  
                group_by(GEOID_TR) %>% 
                summarise(ALLEQ = length(unique(HRA))==1) %>%
                left_join(tr_hra_ids,.,by = 'GEOID_TR') 
        
        # Get the percentage for each count variable
        
        pop_tr <- 
                read_csv(root_file("1-data/3-external/manual/kc-tr/DEC_10_SF1_P1/DEC_10_SF1_P1_with_ann.csv"), 
                         col_types = cols(GEO.id2 = col_character(), 
                                          Geography = col_skip(), 
                                          Id = col_skip(), 
                                          Id2 = col_character()), 
                         skip = 1) %>% 
                select(GEOID_TR = Id2,
                       POP_TR = Total)
        
        hu_tr <- 
                read_csv(root_file('1-data/3-external/manual/kc-tr/DEC_10_SF1_H1/DEC_10_SF1_H1_with_ann.csv'), 
                         col_types = cols(GEO.id2 = col_character(), 
                                          Geography = col_skip(), 
                                          Id = col_skip(), 
                                          Id2 = col_character()), 
                         skip = 1)%>% 
                select(GEOID_TR = Id2,
                       HU_TR = Total)
        
        pophu_tr <- 
                read_csv(root_file('1-data/3-external/manual/kc-tr/DEC_10_SF1_H10/DEC_10_SF1_H10_with_ann.csv'), 
                         col_types = cols(GEO.id2 = col_character(), 
                                          Geography = col_skip(), 
                                          Id = col_skip(), 
                                          Id2 = col_character()), 
                         skip = 1) %>% 
                select(GEOID_TR = Id2,
                       POPHU_TR = Total)
        
        kc_tr_sp@data %<>%
                left_join(.,tr_hra_ids2,by = 'GEOID_TR') %>% 
                left_join(.,pop_tr,by = 'GEOID_TR') %>% 
                left_join(.,hu_tr,by = 'GEOID_TR') %>% 
                left_join(.,pophu_tr,by = 'GEOID_TR') %>% 
                mutate(POP_PCT = round_any(POP/POP_TR,accuracy = .01),
                       HU_PCT = round_any(HU/HU_TR,accuracy = .01),
                       POPHU_PCT = round_any(POPHU/POPHU_TR,accuracy = .01)) %>% 
                select(GEOID_TR,
                       POP,POP_PCT,
                       HU, HU_PCT,
                       POPHU, POPHU_PCT,everything())
        
        kc_tr_sp %>% 
                writeOGR(dsn = root_file('1-data/4-interim/kc-hra-tr-sp.gpkg'),
                         layer = 'kc_hra_tr_sp',
                         driver = 'GPKG',
                         verbose = FALSE,
                         overwrite_layer = TRUE)
        
}

kc_hra_tr_sp <- 
        readOGR(dsn = root_file('1-data/4-interim/kc-hra-tr-sp.gpkg'),
                layer = 'kc_hra_tr_sp',
                verbose = FALSE) %>% 
        spTransform(crs_proj)

```

### Maps


```{r hra-maps}

mypal <- RColorBrewer::brewer.pal(8,name = 'Set2')[-8]

shuffled_hra <- forcats::fct_shuffle(kc_hra_tr_sp$HRA_POP) %>% factor(ordered = T)

pal <- colorFactor(palette = mypal,domain = shuffled_hra)

hra %<>% spTransform(crs_proj)

hra_popup <- paste0(kc_hra_tr_sp$HRA_POP)

show_hra_tr <- function(){
        myLfltGrey(bumpLabels = FALSE,hideControls = FALSE) %>%
        addProviderTiles(providers$CartoDB) %>% 
        addPolygons(data = kc_hra_tr_sp,
                    smoothFactor = 0,
                    weight = 1,
                    color = col2hex("white"),
                    opacity = .85,
                    fillColor = ~pal(shuffled_hra),
                    fillOpacity = .5,
                    group = 'HRA Tract by Pop.',
                    popup = ~paste0(kc_hra_tr_sp$GEOID_TR)) %>% 
        addPolygons(data = hra,
                    smoothFactor = 0,
                    fillOpacity = 0,
                    weight = 2,
                    color = ~pal(factor(hra$HRA2010v2_,levels = levels(shuffled_hra),ordered = TRUE)),
                    opacity = 1,
                    group = 'HRAs',
                    popup = ~paste0(hra$HRA2010v2_)) %>% 
        addLayersControl(overlayGroups = c('HRA Tract by Pop.',
                                           'HRAs'),
                         position = 'topright',options = layersControlOptions(FALSE))
}

# Save an HTML version of the map

if(!file.exists(root_file('3-communication/others/html/hra-tracts.html'))){
        show_hra_tr() %>% 
        saveWidget(file = root_file('3-communication/others/html/hra-tracts.html'),
                           selfcontained = FALSE,
                           libdir = root_file('3-communication/others/html/html_support_files'))
}

show_hra_tr()


```






































