---
df_print: tibble
output:
  html_notebook:
    code_folding: hide
  pdf_document:
    keep_tex: yes
always_allow_html: yes
---

```{r misc-setup, echo = FALSE, warning=FALSE,message=FALSE,comment=FALSE}
library(knitr)
library(rprojroot)
library(tidyverse)
library(rgdal)
library(sp)
library(rgeos)
library(miscgis)
library(tigris)
library(leaflet)
library(ggthemes)
library(magrittr)
library(stringr)
library(downloader)
root <- rprojroot::is_rstudio_project
root_file <- root$make_fix_file()
opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, comment=FALSE)
```

###Seattle Boundary

```{r misc-sea-bound, fig.cap="Seattle\'s geographic boundary"}

if(!file.exists(root_file('1-data/4-interim/seattle.gpkg'))){
        tigris::places(state = "WA") %>%
        tigris::filter_place(place = "Seattle") %>%
        spTransform(CRSobj = crs_proj) %>% 
        writeOGR(dsn = root_file('1-data/4-interim/seattle.gpkg'),
                 layer = "seattle", driver = "GPKG", verbose = FALSE, 
                 overwrite_layer = TRUE)
}

sea <- readOGR(dsn = root_file('1-data/4-interim/seattle.gpkg'),layer = "seattle",verbose = FALSE)

green <- miscgis::miscgis_pals$tableau_cat[["green"]]

myLfltGrey() %>% addPolygons(data = sea,color = green,opacity = 1,fillColor = green,fillOpacity = .5)

```

###King County Subdivision Boundary

```{r misc-sea-ccd, fig.cap="Seattle Subdivision of King County\'s geographic boundary"}

if(!file.exists(root_file('1-data/4-interim/seattle_ccd.gpkg'))){
        tigris::county_subdivisions(state = "53",county = "033") %>% 
                subset(NAME == "Seattle") %>% 
                spTransform(CRSobj = crs_proj) %>% 
                writeOGR(dsn = root_file('1-data/4-interim/seattle_ccd.gpkg'),
                         layer = "seattle", driver = "GPKG", verbose = FALSE, 
                         overwrite_layer = TRUE)
}

sea_ccd <- readOGR(dsn = root_file('1-data/4-interim/seattle_ccd.gpkg'),layer = "seattle",verbose = FALSE) %>% spTransform(crs_proj)

orange <- miscgis::miscgis_pals$tableau_cat[["orange"]]

myLfltGrey() %>% addPolygons(data = sea_ccd,color = orange,opacity = 1,fillColor = orange,fillOpacity = .5)

```

### Tracts in Seattle CCD

In the absence of a straight-forward method for identifying all the census tracts in the Seattle CCD subdivision of King County, it is possible to extract this information from [American Factfinder](https://factfinder.census.gov/faces/nav/jsf/pages/index.xhtml). This [tutorial](https://ask.census.gov/faq.php?id=5000&faqId=1605) describes how to use the American Factfinder interface to extract a list of all "all tracts within (or partially within) a census place"; substituting "county subdivision" for "place" will retrieve the desired results.

```{r misc-tracts-ccd}


if(!file.exists(root_file('1-data/4-interim/tr-ccd-sp.gpkg'))){
        tr_ccd_geoid <- 
        read_csv(
                root_file('1-data/3-external/manual/seattle-ccd/ACS_12_5YR_B01001/ACS_12_5YR_B01001_with_ann.csv'), 
                col_types = cols(Id2 = col_character()), 
                skip = 1) %>% 
        mutate(NEW_GEOID1 = str_sub(Id2,1,5),
               NEW_GEOID2 = str_sub(Id2,16,21),
               GEOID = paste0(NEW_GEOID1,NEW_GEOID2)) %>% 
        select(GEOID) %>% unlist(use.names = F)
        
        # If the WA census tracts objects isn't yet loaded in the global env.,
        # download and load it
        if(!exists('tr')){tr <- tigris::tracts(state = "WA",county = "King", year = 2014) %>% spTransform(CRSobj = crs_proj)}
        
tr_ccd_sp <- subset(tr, GEOID %in% tr_ccd_geoid)

writeOGR(obj = tr_ccd_sp,
         dsn = root_file('1-data/4-interim/tr-ccd-sp.gpkg'),
         layer = 'tr_ccd_sp',
         driver = 'GPKG',
         verbose = FALSE,
         overwrite_layer = TRUE)
}

tr_ccd_sp <- readOGR(dsn = root_file('1-data/4-interim/tr-ccd-sp.gpkg'),
                     layer = 'tr_ccd_sp',
                     verbose = FALSE)
show_tr_ccd_sp <- function(){
        myLfltGrey() %>% 
        addPolygons(data = tr_ccd,weight = .5,color = blue,opacity = 1,fillColor = blue,fillOpacity = .5) %>% 
        addPolygons(data = sea_ccd,color = orange,opacity = 1,fillOpacity = 0)
}
# show_tr_ccd_sp()
```

###Puget Sound Waterbodies

These are useful for "clipping" census geographies whose boundaries extend into waterbodies.

```{r misc-wtr, fig.cap="Puget Sound waterbodies"}

fp_wtr <- root_file('1-data/3-external/NHDMajor.gdb')

# check if the file already exists, if not then download it
if(!file.exists(fp_wtr)){
        
        url <- "ftp://www.ecy.wa.gov/gis_a/inlandWaters/NHD/NHDmajor.gdb.zip" # save the URL for the waterbodies data
        
        temp <- tempfile() # create a temporary file to hold the compressed download
        
        download(url, dest = temp, mode="wb") # download the file
        
        unzip (temp, exdir = root_file('1-data/3-external/')) # extract the ESRI geodatabase file to a project folder
}

wtr_sp <-
        suppressWarnings(readOGR(dsn = fp_wtr,      # create a waterbodies shape
                layer = "NHD_MajorWaterbodies",verbose = FALSE,pointDropZ = TRUE)) %>%
        gBuffer(byid=TRUE, width=0) %>% # clean up self-intersecting polygons
        spTransform(CRSobj = crs_proj)  # transform the projection to match the project projection
        

intersect <- gIntersects(sea_ccd,wtr_sp,byid = TRUE) %>% as.vector()

wtr_sp <- wtr_sp[intersect,] %>%
        spTransform(CRSobj = crs_proj)

blue <- miscgis::miscgis_pals$tableau_cat[["blue"]]

if(!file.exists(root_file('1-data/4-interim/wtr_sp.gpkg'))){
        writeOGR(obj = wtr_sp, dsn = root_file('1-data/4-interim/wtr_sp.gpkg'), 
          layer = "wtr_sp", driver = "GPKG", verbose = FALSE, 
          overwrite_layer = TRUE)
}

show_wtr <- function(){
        myLfltGrey() %>% 
                addPolygons(data = wtr_sp,color = blue, opacity = 1, 
                            weight = .5, fillColor = blue,fillOpacity = .5)        
}

show_wtr()

```
