---
df_print: tibble
output:
  html_notebook:
    code_folding: hide
  pdf_document:
    keep_tex: yes
always_allow_html: yes
---

```{r poc-setup, echo = FALSE, warning=FALSE,message=FALSE,comment=FALSE}
library(plyr)
library(knitr)
library(rprojroot)
library(tidyverse)
library(rgdal)
library(sp)
library(rgeos)
library(miscgis)
library(tigris)
library(leaflet)
library(ggthemes)
library(magrittr)
library(stringr)
library(downloader)
library(webshot)
library(htmltools)
library(gplots)
library(ggmap)
library(shiny)
library(htmlwidgets)
library(readxl)
library(acs)
library(RColorBrewer)
root <- rprojroot::is_rstudio_project
root_file <- root$make_fix_file()
opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, comment=FALSE)
```

###Introduction
This datset comes from the [American Community Survey](https://www.census.gov/programs-surveys/acs/). General information about the structure of this dataset can be found [here](https://factfinder.census.gov/faces/affhelp/jsf/pages/metadata.xhtml?lang=en&type=table&id=table.en.ACS_14_5YR_B03002#main_content).

This dataset is used to create the indicator "Percent People of Color", which is defined as the percentage of the population that is a race other than non-Hispanic White.

###Metadata

```{r poc-metadata, fig.cap='Metadata | People of Color'}

tibble(
        "Program" = "American Community Survey",
        "Data Set" = "2010-2014 American Community Survey 5-Year Estimates",
        "Table" = "B03002 - HISPANIC OR LATINO ORIGIN BY RACE"
           ) %>% knitr::kable(caption = 'Metadata | People of Color')

```

```{r poc-colnames}

if(!file.exists(root_file('./1-data/3-external/poc-pretty20.rds'))){
        tbl <- "B03002"  # census table code
        geog <- geo.make(state = "WA",county = "King",tract = "*")
        
        #Coded column names
        acs.fetch(endyear = "2014",
                  span = 5,
                  geography = geog,
                  table.number = tbl) %>% readr::write_rds(root_file('./1-data/3-external/poc-2014.rds'))
        
        # "Pretty" column names
        acs.fetch(endyear = "2014",
                  span = 5,
                  geography = geog,
                  table.number = tbl,
                  col.names = "pretty") %>% readr::write_rds("./1_raw/poc_pretty.rds")
}


poc_acs <- read_rds("./1_raw/poc.rds")
poc_prty_acs <- read_rds("./1_raw/poc_pretty.rds")

knitr::kable(tibble("Abbr." = acs.colnames(poc_acs),
                    "Full Column Name" = acs.colnames(poc_prty_acs)),
             row.names = 1:ncol(poc_acs))
```


###Raw Data
The data is structured in a special format that retains important metadata about the variable, including the standard error values needed to calculate confidence intervals.
```{r,echo= FALSE}
head(poc_prty_acs[,1])
```




###New Column: People of Color
```{r}
poc_acs2 <- poc_acs[, "B03002_001"] - poc_acs[, "B03002_003"]
                                
acs.colnames(poc_acs2) <- "POC"

poc_acs2 %<>% cbind(poc_acs, .)

poc_acs3 <- 
        apply(
                X = poc_acs2[, 22],
                MARGIN = 1,
                FUN = divide.acs,
                denominator = poc_acs2[, 1],
                method = "proportion",
                verbose = FALSE
        )

acs.colnames(poc_acs3) <- "POC_PCT"

readr::write_rds(poc_acs3,"./2_intermediate/poc_acs.rds")

head(poc_acs3[])

```

###Conversion to a Dataframe

```{r}

poc_acs <- readr::read_rds("./2_intermediate/poc_acs.rds")

poc_df <- 
        data.frame(
                geography(poc_acs)["tract"], 
                estimate(poc_acs), 
                1.645 * standard.error(poc_acs)) %>% 
        `colnames<-`(., c("GEOID6", "POC_PCT_EST","POC_PCT_MOE")) %>% 
        mutate(UPPER = POC_PCT_EST + POC_PCT_MOE, 
               LOWER = POC_PCT_EST - POC_PCT_MOE, 
               UPPER = if_else(UPPER > 1, 1, UPPER), 
               LOWER = if_else(LOWER < 0, 0, LOWER))

readr::write_rds(poc_df,"./2_intermediate/poc_df.rds")
as_tibble(poc_df)

```

###Mapping

```{r, echo=FALSE,message=FALSE,fig.height=1.25, fig.width=1,  fig.show='hold', dpi=150,fig.cap='Rainier Valley, White Center, and SeaTac/Tukwila'}




```







