---
df_print: tibble
output:
  html_notebook: default
  pdf_document:
    keep_tex: yes
always_allow_html: yes
---

```{r res-val-setup, echo = FALSE, warning=FALSE,message=FALSE}
library(knitr)
library(rprojroot)
library(tidyverse)
library(rgdal)
library(sp)
library(rgeos)
library(miscgis)
library(tigris)
library(leaflet)
library(ggthemes)
library(magrittr)
library(downloader)
root <- rprojroot::is_rstudio_project
root_file <- root$make_fix_file()
opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, comment = FALSE)
options("readr.num_columns" = 0)
```

### Introduduction

This dataset is a collection of property and building appraised values.

```{r res-val-dl}

if(!file.exists(root_file('1-data/3-external/EXTR_RealPropApplHist_V.csv'))){
        
        # direct URL to the file download
        url <- 'http://aqua.kingcounty.gov/extranet/assessor/Real%20Property%20Appraisal%20History.zip' 
        
        temp <- tempfile() # create a temporary file to hold the compressed download
        
        download(url, dest = temp, mode='wb') # download the file (~260MB)
        
        unzip (temp, exdir = root_file('1-data/3-external/')) # extract the file to the project folder
}

if(!exists('res_val_csv')){
        res_val_csv <- read_csv(
        root_file("1-data/3-external/EXTR_RealPropApplHist_V.csv"),
        col_types = cols(PostDate = col_datetime(format = "%Y-%m-%d %H:%M:%S"),
                         RollYr = col_date(format = "%Y"),
                         SelectDate = col_datetime(format = "%Y-%m-%d %H:%M:%S"),
                         UpdateDate = col_datetime(format = "%Y-%m-%d %H:%M:%S")),
        progress = FALSE)
}

# res_val_csv %>% head %>% as_tibble()

```

### Data Transformation
The following data transformations need to be completed in order to for this data to be useful for this project:

  * exclude parcels outside the Seattle CCD
  * exclude non-residential parcels
  * exclude years outside the observations periods (2005, 2010, and 2015)
  * exclude columns with information not relevant to this project (e.g., `SelectMethod`)
  * combine the values of the land and the structure(s) on the land (`LandVal` and `ImprsVal`, respectively)
  * adjust dollar amount for inflation based on 2016 US Dollar value
  * associate parcels with census tracts so that summary statistics can be calculated (e.g. median)


```{r res-val-rds, fig.cap='Parcel dataset attributes (5 randomly selected rows)'}

# Given the size of this dataset (~13M records, 1.8GB), it's useful to perform the simplest 
# filters first, then save and reload the reduced-size subset (~2M records, 89MB).

res_val_orig <- res_val_csv

colnames(res_val_orig) %<>% toupper()

if(!file.exists(root_file('1-data/4-interim/res-val-sub1.Rds'))){
        res_val_orig %>% 
        select(MAJOR:NEWDOLLARS,-REVALORMAINT) %>% 
        filter(ROLLYR %in% lubridate::as_date(c("2005-01-01","2010-01-01","2015-01-01"))) %>% 
        write_rds(root_file('1-data/4-interim/res-val-sub1.Rds'))
}

res_val_sub1 <- read_rds(path = root_file('1-data/4-interim/res-val-sub1.Rds'))

res_val_sub2 <- 
        res_val_sub1 %>% 
        mutate(PIN = paste0(MAJOR,MINOR)) %>% 
        select(PIN,everything())

set.seed(1)
res_val_sub2 %>% sample_n(5) %>% kable(caption = 'Parcel dataset attributes (5 randomly selected rows)',booktabs = TRUE)

```



