---
df_print: tibble
output:
  html_notebook:
    code_folding: hide
  pdf_document:
    keep_tex: yes
always_allow_html: yes
---

```{r tracts-setup, echo = FALSE, warning=FALSE,message=FALSE,comment=FALSE}
library(knitr)
library(rprojroot)
library(tidyverse)
library(rgdal)
library(sp)
library(rgeos)
library(miscgis)
library(tigris)
library(leaflet)
library(ggthemes)
library(magrittr)
library(stringr)
library(downloader)
library(webshot)
library(htmltools)
library(gplots)
library(ggmap)
library(shiny)
library(htmlwidgets)
library(readxl)
root <- rprojroot::is_rstudio_project
root_file <- root$make_fix_file()
opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, comment=FALSE)
```



### Introduction
The primary geographic unit of this assessment is the [census tract](https://www.census.gov/geo/reference/gtc/gtc_ct.html). As is the case with many communities, the census geographies do not coincide exactly with the formal geographic boundary of the study's three sites, and should be considered as _spatial approximations_ of these communities.

Listed below are the geographic identifiers of the census tracts that approximate each site.


```{r tracts-intro, results="asis"}

# Rainier Valley
rv <- c("53033010001", "53033010300", "53033010401", "53033011001", "53033011002", "53033011101","53033011102", "53033011700", "53033011800", "53033011900")

# White Center
wc <- c("53033026600","53033026700","53033026500","53033026801","53033026802","53033027000")

# SeaTac/Tukwila

stc_hus <- read_csv(root_file('1-data/3-external/manual/seatac-tr/DEC_10_SF1_H1/DEC_10_SF1_H1_with_ann.csv'), 
                    col_types = cols(Id2 = col_character()), skip = 1)

tuk_hus <- read_csv(root_file('1-data/3-external/manual/tukwila-tr/DEC_10_SF1_H1/DEC_10_SF1_H1_with_ann.csv'), 
    col_types = cols(Id2 = col_character()), 
    skip = 1)

stc <-  paste(substr(x = stc_hus$Id2,1,2),substr(x = stc_hus$Id2,8,16),sep = "")

tuk <-  paste(substr(x = tuk_hus$Id2,1,2),substr(x = tuk_hus$Id2,8,16),sep = "")

stctuk <- c(stc,tuk)

df <- miscgis::cbind_fill(rv,wc,stctuk) %>% 
        as_tibble() %>% 
        `colnames<-`(c("Rainier Valley",
                       "White Center",
                       "SeaTac/Tukwila"))

knitr::kable(df,caption = "Census Tract GEOIDs")
```

### Spatial Polygons

#### With Waterbodies
Left to right: Rainier Valley, White Center, SeaTac/Tukwila
<br></br>

```{r tracts-with-wtr, fig.height=1.25, fig.width=1,  fig.show='hold', dpi=150,fig.cap='Rainier Valley, White Center, and SeaTac/Tukwila'}

# library(htmltools)
# library(htmlwidgets)
# library(leaflet)
# library(gplots)
# library(ggmap)
# library(rgeos)
# library(sp)
# library(magrittr)
# library(rgdal)

if(!file.exists(root_file('1-data/3-external/tr_kc.gpkg'))){
        tigris::tracts(state = "WA",county = "King", year = 2014) %>% 
        spTransform(CRSobj = crs_proj) %>% 
        writeOGR(dsn = root_file('1-data/3-external/tr_kc.gpkg'),
                 layer = 'tr_kc_sp',
                 driver = 'GPKG',
                 verbose = FALSE,  
                 overwrite_layer = TRUE)
}


tr <- readOGR(dsn = root_file('1-data/3-external/tr_kc.gpkg'),
              layer = 'tr_kc_sp',
              verbose = FALSE,
              stringsAsFactors = FALSE) %>% spTransform(crs_proj)

if(!file.exists(root_file("1-data/4-interim/tr_sp_withwater.gpkg"))){
        
tr_sel <- tr[tr$GEOID %in% c(rv,wc,stctuk),] %>% spTransform(CRSobj = crs_proj)

tr_sp <- tr_sel

tr_sp@data %<>% mutate(COO_SITE = case_when(.$GEOID %in% rv ~ "Rainier Valley",
                                            .$GEOID %in% wc ~ "White Center",
                                            .$GEOID %in% stctuk ~ "SeaTac/Tukwila"))
        
writeOGR(obj = tr_sp,
         dsn = root_file("1-data/4-interim/tr_sp_withwater.gpkg"),
         layer = "tr_sp_withwater",
         driver =  "GPKG",
         verbose = FALSE,
         overwrite_layer = TRUE)
}

tr_sp <- readOGR(dsn = root_file("1-data/4-interim/tr_sp_withwater.gpkg"),
         layer = "tr_sp_withwater",
         stringsAsFactors = FALSE,
         verbose = FALSE)


cntr <- tr_sp %>% gCentroid(byid = TRUE) %>% gCentroid()

blue <- miscgis::miscgis_pals$tableau_cat[["blue"]]
green <- miscgis::miscgis_pals$tableau_cat[["green"]]
red <- miscgis::miscgis_pals$tableau_cat[["red"]]
teal <- miscgis::miscgis_pals$tableau_cat[["teal"]]
brown <- miscgis::miscgis_pals$tableau_cat[["brown"]]


# Rainier Valley
        
rv_map <- miscgis::myLfltGrey() %>%
        addPolygons(data = tr_sp[tr_sp$COO_SITE == "Rainier Valley",],
                    smoothFactor = 0,
                    color = col2hex("white"),
                    weight = .5,
                    opacity = 1,
                    fillColor = blue,
                    fillOpacity = .65
        ) %>% 
        setView(cntr$x,cntr$y,zoom = 9) %>% 
        miscgis::styleWidget(style = "float:left;margin:1px")



# White Center


wc_map <- miscgis::myLfltGrey() %>%
        addPolygons(data = tr_sp[tr_sp$COO_SITE == "White Center",],
                    smoothFactor = 0,
                    color = col2hex("white"),
                    weight = .5,
                    opacity = 1,
                    fillColor = green,
                    fillOpacity = .65
        ) %>% 
        setView(cntr$x,cntr$y,zoom = 9) %>% 
        miscgis::styleWidget(style = "float:left;margin:1px")

# SeaTac/Tukwila

stctuk_map <- miscgis::myLfltGrey() %>%
        addPolygons(data = tr_sp[tr_sp$COO_SITE == "SeaTac/Tukwila",],
                    smoothFactor = 0,
                    color = col2hex("white"),
                    weight = .5,
                    opacity = 1,
                    fillColor = red,
                    fillOpacity = .65
        ) %>% 
        setView(cntr$x,cntr$y,zoom = 9) %>% 
        miscgis::styleWidget(style = "float:none;margin:1px")

rv_map
wc_map
stctuk_map


```


#### Without Waterbodies
Left to right: Rainier Valley, White Center, SeaTac/Tukwila
<br></br>

```{r tracts-wo-wtr, fig.height=1.25, fig.width=1,  fig.show='hold', dpi=150,fig.cap='Rainier Valley, White Center, and SeaTac/Tukwila'}


if(!file.exists(root_file('1-data/4-interim/tr_trim_sp.gpkg'))){
        
        wtr <- readOGR(dsn = root_file('1-data/4-interim/wtr_sp.gpkg'), 
          layer = "wtr_sp",verbose = FALSE) %>% gUnaryUnion(.) %>% spTransform(CRSobj = crs_proj)

spClip(tr_sp,wtr) %>% 
        subset(!duplicated(GEOID)) %>% 
        writeOGR(dsn = root_file('1-data/4-interim/tr_trim_sp.gpkg'),
         layer = "tr_trim_sp",driver = "GPKG",
         verbose = FALSE,
         overwrite_layer = TRUE)
}

tr_trim_sp <- readOGR(dsn = root_file('1-data/4-interim/tr_trim_sp.gpkg'),
         layer = "tr_trim_sp",
         verbose = FALSE,
         stringsAsFactors = FALSE)

# Individiaul Maps

rv_trim <- tr_trim_sp %>% subset(COO_SITE == "Rainier Valley")

wc_trim <- tr_trim_sp %>% subset(COO_SITE == "White Center")

stctuk_trim <- tr_trim_sp %>% subset(COO_SITE == "SeaTac/Tukwila")


# Rainier Valley
        
rv_trim_map <- miscgis::myLfltGrey() %>%
        addPolygons(data = rv_trim,
                    smoothFactor = 0,
                    color = col2hex("white"),
                    weight = .5,
                    opacity = 1,
                    fillColor = blue,
                    fillOpacity = .65
        ) %>% 
        setView(cntr$x,cntr$y,zoom = 9) %>% 
        miscgis::styleWidget(style = "float:left;margin:1px")



# White Center

wc_trim_map <- miscgis::myLfltGrey() %>%
        addPolygons(data = wc_trim,
                    smoothFactor = 0,
                    color = col2hex("white"),
                    weight = .5,
                    opacity = 1,
                    fillColor = green,
                    fillOpacity = .65
        ) %>% 
        setView(cntr$x,cntr$y,zoom = 9) %>% 
        miscgis::styleWidget(style = "float:left;margin:1px")

# SeaTac/Tukwila

stctuk_trim_map <- miscgis::myLfltGrey() %>%
        addPolygons(data = stctuk_trim,
                    smoothFactor = 0,
                    color = col2hex("white"),
                    weight = .5,
                    opacity = 1,
                    fillColor = red,
                    fillOpacity = .65
        ) %>% 
        setView(cntr$x,cntr$y,zoom = 9) %>% 
        miscgis::styleWidget(style = "float:none;margin:1px")

rv_trim_map

wc_trim_map

stctuk_trim_map



```


#### Health Reporting Areas (Disaggregated by Tract)
```{r tracts-combined, dpi = 150}
# Combined map 

# if(!exists('tr')){
#         tr <- readOGR(dsn = root_file('1-data/3-external/tr_kc.gpkg'),
#               layer = 'tr_kc_sp',
#               verbose = FALSE,
#               stringsAsFactors = FALSE) %>% spTransform(crs_proj)
# }
# 
# pts <- gCentroid(tr,byid = TRUE) %>% spTransform(crs_proj) %>% 
#         SpatialPointsDataFrame(data = tr@data,match.ID = FALSE)
# 
# if(!exists('tr_trim_sp')){
# tr_trim_sp <- readOGR(dsn = root_file('1-data/4-interim/tr_trim_sp.gpkg'),layer = 'tr_trim_sp',verbose = FALSE,stringsAsFactors = FALSE)
# }
# 
# popups <- paste0(tr_trim_sp@data$GEOID)
# 
# hra <- readOGR(dsn = root_file('1-data/3-external/manual/HRA_2010Block_Clip/'),
#                layer = "HRA_2010Block_Clip",stringsAsFactors = FALSE,verbose = FALSE) %>% spTransform(crs_proj)
# 
# rv_hra_id <- c("SE Seattle",
#                "Beacon/Gtown/S.Park")
# 
# wc_hra_id <- c("North Highline")
# 
# stctuk_hra_id <- c("SeaTac/Tukwila")
# 
# 
# hra@data %<>% mutate(HRA_ID = case_when(.$HRA2010v2_ %in% rv_hra_id ~ "Rainier Valley",
#                                         .$HRA2010v2_ %in% wc_hra_id ~ "White Center",
#                                         .$HRA2010v2_ %in% stctuk_hra_id ~ "SeaTac/Tukwila"),
#                      HRA_ID_FCT = factor(HRA_ID,
#                                    levels = c("Rainier Valley",
#                                 "White Center",
#                                 "SeaTac/Tukwila"),ordered = TRUE)
#                      )
# 
# hra_coo <- hra %>% subset(!is.na(HRA_ID))
# 
# popup_hra <- paste0(hra_coo$HRA2010v2_)
# 
# # rv_tr_hra <- c(53033010100,53033010300,53033010200,53033011102,53033011101,53033011800,53033011900,53033026001) %>% as.character()
# # 
# # wc_tr_hra <- c(53033009300,53033010002,53033010001,53033010402,53033010401,53033011002,53033011001,53033010900,53033011200,53033011700,53033026400,53033026500,53033026600,53033026700,53033027000,53033026802,53033026801) %>% as.character()
# # 
# # stctuk_tr_hra <- c(53033026300,53033026400,53033027200,53033027100,53033026200,53033028200,53033028100,53033027300,53033028000,53033028500,53033028402,53033028300,53033028403,53033028802,53033028700,53033028801,53033028500,53033029101) %>% 
# #         as.character()
# 
# site_hra_tr <- read_excel(root_file('/1-data/3-external/manual/Definition_GeographicUnit_COO_20160913-tm-edit.xlsx'), 
#                           col_types = c("text", "text", "numeric", "text", "text", "text", "text"))
# colnames(site_hra_tr) %<>% toupper()
# 
# site_hra_tr %<>% 
#         select(GEOID,SITES) %>% 
#         mutate(SITES2 = SITES) %>% 
#         group_by(GEOID) %>% 
#         mutate(SITES2 = nth(SITES,2),
#                SITES2 = ifelse(SITES == SITES2,NA_character_,SITES2)) %>% 
#         transmute(SITES = ifelse(is.na(SITES2),SITES,paste(SITES,SITES2,sep = ":")),
#                   SITES_FCT = factor(SITES,
#                      levels = c("Rainier Valley",
#                                 "Rainier Valley:SeaTac/Tukwila",
#                                 "White Center",
#                                 "White Center:SeaTac/Tukwila",
#                                 "SeaTac/Tukwila"),ordered = TRUE))
# 
# 
# site_names <- site_hra_tr$SITES_FCT %>% levels() %>% as.character()
# 
# hra_geoids_lst <- split(site_hra_tr, site_hra_tr$SITES_FCT) %>% 
#         map(.f = ~.['GEOID']) %>% flatten() %>% 
#         `names<-`(site_names)
# 
# tr_trim_sp@data %<>% mutate(HRA_COO_SITE = case_when(
#         .$GEOID %in% unlist(hra_geoids_lst["Rainier Valley"], use.names = FALSE) ~ "Rainier Valley",
#         .$GEOID %in% unlist(hra_geoids_lst["Rainier Valley:SeaTac/Tukwila"], use.names = FALSE) ~ "Rainier Valley:SeaTac/Tukwila",
#         .$GEOID %in% unlist(hra_geoids_lst["White Center"], use.names = FALSE) ~ "White Center",
#         .$GEOID %in% unlist(hra_geoids_lst["White Center:SeaTac/Tukwila"], use.names = FALSE) ~ "White Center:SeaTac/Tukwila",
#         .$GEOID %in% unlist(hra_geoids_lst["SeaTac/Tukwila"], use.names = FALSE) ~ "SeaTac/Tukwila"),
#         HRA_COO_SITE_FCT = factor(HRA_COO_SITE,
#                      levels = c("Rainier Valley",
#                                 "Rainier Valley:SeaTac/Tukwila",
#                                 "White Center",
#                                 "White Center:SeaTac/Tukwila",
#                                 "SeaTac/Tukwila"),ordered = TRUE)
#                             )
# 
# tr_hra <- tr
# 
# tr_hra@data %<>% mutate(HRA_COO_SITE = case_when(
#         .$GEOID %in% unlist(hra_geoids_lst["Rainier Valley"], use.names = FALSE) ~ "Rainier Valley",
#         .$GEOID %in% unlist(hra_geoids_lst["Rainier Valley:SeaTac/Tukwila"], use.names = FALSE) ~ "Rainier Valley:SeaTac/Tukwila",
#         .$GEOID %in% unlist(hra_geoids_lst["White Center"], use.names = FALSE) ~ "White Center",
#         .$GEOID %in% unlist(hra_geoids_lst["White Center:SeaTac/Tukwila"], use.names = FALSE) ~ "White Center:SeaTac/Tukwila",
#         .$GEOID %in% unlist(hra_geoids_lst["SeaTac/Tukwila"], use.names = FALSE) ~ "SeaTac/Tukwila"),
#         HRA_COO_SITE_FCT = factor(HRA_COO_SITE,
#                      levels = c("Rainier Valley",
#                                 "Rainier Valley:SeaTac/Tukwila",
#                                 "White Center",
#                                 "White Center:SeaTac/Tukwila",
#                                 "SeaTac/Tukwila"),ordered = TRUE)
#                             )
# 
# if(!exists('wtr_sp')){
# wtr_sp <- readOGR(dsn = root_file('1-data/4-interim/wtr_sp.gpkg'),layer = 'wtr_sp',verbose = FALSE,stringsAsFactors = FALSE) %>% spTransform(crs_proj)
# }
# 
# tr_hra %<>% subset(!is.na(HRA_COO_SITE)) %>% spClip(wtr_sp)
# 
# # Color palettes
# 
# pal_rgb_3 <- miscgis_pals$tableau_cat[c("blue","green","red")] %>% unlist %>% palette()
# pal_rgb_5 <- miscgis_pals$tableau_cat[c("blue","teal","green","brown","red")] %>% unlist %>% palette()
# 
# pal_3 <- colorFactor(pal_rgb_3,
#                    domain = factor(tr_trim_sp$COO_SITE,
#                                    levels = c("Rainier Valley",
#                                 "White Center",
#                                 "SeaTac/Tukwila"),ordered = TRUE))
# pal_5 <- colorFactor(pal_rgb_5,domain = tr_hra$HRA_COO_SITE_FCT)
# 
# 
# if(!file.exists(root_file('3-commmunication/others/html/tr_coo_sites.html'))){
#         leaflet() %>%
#                 addProviderTiles(providers$CartoDB.Positron) %>% 
#                 addPolygons(data = tr_trim_sp,
#                             smoothFactor = 0,
#                             color = col2hex("white"),
#                             weight = .5,
#                             opacity = 1,
#                             fillColor = ~pal_3(COO_SITE),
#                             fillOpacity = .65,
#                             group = "COO Sites - Guess") %>% 
#                 addPolygons(data = hra_coo,
#                             smoothFactor = 0,
#                             color = ~pal_3(HRA_ID_FCT),
#                             weight = 2.5,
#                             opacity = 1,
#                             fillOpacity = 0,
#                             popup = popup_hra,
#                             group = "HRA Boundaries") %>% 
#                 addPolygons(data = tr_hra,
#                             smoothFactor = 0,
#                             color = col2hex("white"),
#                             weight = .5,
#                             opacity = 1,
#                             fillColor = ~pal_5(HRA_COO_SITE_FCT),
#                             fillOpacity = .65,
#                             group = "COO Sites - HRA") %>%
#                 addPolygons(data = tr,
#                             smoothFactor = 0,
#                             weight = .75,
#                             color = col2hex("grey50"),
#                             opacity = .75,
#                             fillOpacity = 0,
#                             group = "All Tracts", popup = popups) %>%
#                 # addPolygons(data = hra,
#                 #             smoothFactor = 0,
#                 #             color = col2hex("white"),
#                 #             weight = .5,
#                 #             opacity = 1,
#                 #             fillColor = ~pal(COO_SITE),
#                 #             fillOpacity = .65,
#                 #             group = "HRA") %>% 
#                 setView(cntr$x,cntr$y,zoom = 11) %>% 
#                 addLayersControl(overlayGroups = c("COO Sites - HRA",
#                                                    "COO Sites - Guess",
#                                                    "HRA Boundaries",
#                                                    "All Tracts")) %>% 
#                 addLegend(title = "COO Sites",
#                           position = "topright",
#                           pal = pal_5,
#                           values = tr_trim_sp$HRA_COO_SITE_FCT) %>% 
#                 saveWidget(file = root_file('3-communication/others/html/tr_coo_sites.html'),
#                            selfcontained = FALSE,
#                            libdir = root_file('3-communication/others/html/html_support_files'))
# }

```

