---
title: "sp × tidyverse test"
author: Tiernan Martin
output: html_notebook
date: November 02, 2016
---

This is a set of tests related to a [Github issue](https://github.com/edzer/sfr/issues/42) for the `sf` package. 

#### Setup
```{r setup}
library(dplyr)
library(sf)
library(RColorBrewer)

demo(nc, ask = FALSE, echo = FALSE)

nc_grp <- 
        nc %>%  
        mutate(GRP = ntile(BIR74,4))

nc_aggr <- 
        nc_grp %>% 
        group_by(GRP) %>% 
        summarise_all(first)

# This works, but I'm having trouble with other examples of the same process
nc_both <- 
        nc %>%  
        mutate(GRP = ntile(BIR74,4)) %>% 
        group_by(GRP) %>% 
        summarise_all(first) %>% 
        rbind(nc_grp)

```

#### Loss of CRS
```{r crs}
all.equal(st_crs(nc),
          st_crs(nc_aggr))
```


#### Unexpected Grouping
```{r groups}
par(mfrow=c(2,2),mar = rep(0,4)) 

cols <- brewer.pal(4, "Spectral")

plot(nc, col = "transparent", border = "black")
plot(nc_aggr, col = "white", border = cols[1])
plot(nc_grp, col = cols[nc_grp$GRP], border = cols[nc_grp$GRP])
plot(nc_grp, col = cols[nc_grp$GRP], border = cols[nc_grp$GRP])
plot(nc_aggr, col = "transparent", border = "white", add = TRUE)
```
Upper-left is the original `nc` object. 

Upper-right is the `nc_aggr`, which has been tranformed with `summarise()` which aggregates some counties together but not what I expected (also notice that the lack of CRS is visually apparent).

Lower-left is `nc_grp` colored by the group (`nc_grp$GRP`), which shows how I expected the aggregated geometries to look.

Lower-right is `nc_grp` overlayed by `nc_aggr` (notice that the white border lines do not align with the fill colors)


#### Geom Dropping
```{r}
par(mfrow=c(2,1),mar = rep(0,4)) 

plot(nc_grp, col = cols[nc_grp$GRP], border = cols[nc_grp$GRP])
plot(nc_grp, col = cols[nc_grp$GRP], border = cols[nc_grp$GRP])
plot(nc_aggr, col = "grey", border = "grey", add = TRUE)
```
Notice that the southern-most county in `nc_grp` (lime green) is dropped from `nc_aggr` (grey)
