--- 
title: Communities of Opportunity Displacement Risk Assessment
author: Tiernan Martin, [Futurewise](http://www.futurewisewa.org/)
date: Fall 2016
subtitle: A User Guide
df_print: tibble
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
description: "This is a report."
---

# Preface {-}

This is a work in-progress.

<!--chapter:end:index.Rmd-->

# (APPENDIX) Appendices {-}

```{r data-methods-setup, echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
library(rprojroot)
library(tidyverse)
root <- rprojroot::is_rstudio_project
root_file <- root$make_fix_file()
opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
options("readr.num_columns" = 0)
```

# Data Methods

## Miscellaneous
```{r data-methods-misc}
misc_fp <- root_file('1-data/1-notebooks/01-misc.Rmd')
```

`r knit_child(misc_fp)`

## Census Geographies: Spatial Data

```{r data-methods-hra-census-acs}

census_geos_fp <- root_file('1-data/1-notebooks/02-census-geos.Rmd')
```

`r knit_child(census_geos_fp)`


## Census Geographies: ACS Geo Sets

```{r data-methods-census-acs}
census_acs_fp <- root_file('1-data/1-notebooks/02-census-acs.Rmd')
```

`r knit_child(census_acs_fp)`

## King County Parcels

```{r data-methods-kc-parcels}
kc_parcels_fp <- root_file('1-data/1-notebooks/kc-parcels.Rmd')
```

`r knit_child(kc_parcels_fp)`

## Vulnerability

```{r data-methods-vuln}
vuln_fp <- root_file('1-data/1-notebooks/03-vulnerability.Rmd')
```

`r knit_child(vuln_fp)`

## Demographic Change

```{r data-methods-demo-chng}

demo_chng_intro_fp <- root_file('1-data/1-notebooks/03-demo-change-intro.Rmd')

white_fp <- root_file('1-data/1-notebooks/03-demo-change-white.Rmd')

bach_fp <- root_file('1-data/1-notebooks/03-demo-change-bach.Rmd')

own_fp <- root_file('1-data/1-notebooks/03-demo-change-own.Rmd')

income_fp <- root_file('1-data/1-notebooks/03-demo-change-inc.Rmd')

```

`r knit_child(demo_chng_intro_fp)`
`r knit_child(white_fp)`
`r knit_child(bach_fp)`
`r knit_child(own_fp)`
`r knit_child(income_fp)`


<!--chapter:end:data-methods.Rmd-->

