# (APPENDIX) Appendices {-}

```{r data-methods-setup, echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
library(rprojroot)
library(tidyverse)
root <- rprojroot::is_rstudio_project
root_file <- root$make_fix_file()
opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
options("readr.num_columns" = 0)
```

# Data Methods

## Miscellaneous
```{r data-methods-misc}
misc_fp <- root_file('1-data/1-notebooks/01-misc.Rmd')
```

`r knit_child(misc_fp)`

## Census Data

```{r data-methods-census}

census_acs_fp <- root_file('1-data/1-notebooks/02-census-acs.Rmd')

census_acs_comms_fp <- root_file('1-data/1-notebooks/02-census-acs-communities.Rmd')

```

`r knit_child(census_acs_fp)`
<br>
`r knit_child(census_acs_comms_fp)`

## King County Assessor Data

```{r data-methods-kc-parcels}

kc_parcels_intro_fp <- root_file('1-data/1-notebooks/02-king-county-parcels-intro.Rmd')

kc_parcels_fp <- root_file('1-data/1-notebooks/02-king-county-parcels.Rmd')
```

`r knit_child(kc_parcels_intro_fp)`
<br>
`r knit_child(kc_parcels_fp)`


```{r data-methods-inds-setup}

# Setup scripts

make_acs_inds_fp <- root_file('1-data/1-notebooks/03-make-acs-inds.Rmd')

make_acs_maps_fp <- root_file('1-data/1-notebooks/03-make-acs-maps.Rmd')

make_acs_sparklines_fp <- root_file('./1-data/4-interim/acs-sparkline-tbl.rds')

make_hous_geoms_fp <- root_file('1-data/1-notebooks/03-make-prcl-geoms.Rmd')

make_hous_meta_fp <- root_file('1-data/1-notebooks/03-make-prcl-meta.Rmd')

make_hous_by_geoms_fp <- root_file('1-data/1-notebooks/03-make-prcl-aggr-by-geom.Rmd')

make_hous_val_inds_fp <- root_file('1-data/1-notebooks/03-make-hous-val-inds.Rmd')

```

`r knit_child(make_acs_inds_fp)`
`r knit_child(make_acs_maps_fp)`
`r knit_child(make_acs_sparklines_fp)`
`r knit_child(make_hous_geoms_fp)`
`r knit_child(make_hous_meta_fp)`
`r knit_child(make_hous_by_geoms_fp)`

## Vulnerability Indicators   
```{r data-methods-vuln-inds}
# Visualization scripts

vuln_intro_fp <- root_file('1-data/1-notebooks/03-vuln-intro.Rmd')

vuln_poc_fp <- root_file('1-data/1-notebooks/03-vuln-poc.Rmd')

vuln_rent_fp <- root_file('1-data/1-notebooks/03-vuln-rent.Rmd')

vuln_lt_bach_fp <- root_file('1-data/1-notebooks/03-vuln-lt-bach.Rmd')

vuln_low_inc_fp <- root_file('1-data/1-notebooks/03-vuln-low-inc.Rmd')


```

`r knit_child(vuln_intro_fp)`
<br>
<hr>
`r knit_child(vuln_poc_fp)`
<br>
<hr>
`r knit_child(vuln_lt_bach_fp)`
<br>
<hr>
`r knit_child(vuln_low_inc_fp)`
<br>
<hr>
`r knit_child(vuln_rent_fp)`
<br>


## Demographic Change Indicators   
```{r data-methods-demo-chng-inds}

demo_chng_intro_fp <- root_file('1-data/1-notebooks/03-demo-change-intro.Rmd')

demo_chng_white_fp <- root_file('1-data/1-notebooks/03-demo-change-white.Rmd')

demo_chng_own_fp <- root_file('1-data/1-notebooks/03-demo-change-own.Rmd')

demo_chng_bach_fp <- root_file('1-data/1-notebooks/03-demo-change-bach.Rmd')

demo_chng_income_fp <- root_file('1-data/1-notebooks/03-demo-change-inc.Rmd')

```

`r knit_child(demo_chng_intro_fp)`
<br>
<hr>
`r knit_child(demo_chng_white_fp)`
<br>
<hr>
`r knit_child(demo_chng_bach_fp)`
<br>
<hr>
`r knit_child(demo_chng_income_fp)`
<br>
<hr>
`r knit_child(demo_chng_own_fp)`
<br>


## Housing Market Indicators
```{r data-methods-hous-inds}
hous_intro_fp <- root_file('1-data/1-notebooks/03-housing-market-intro.Rmd')

# hous_ratio_fp
# 
# hous_rate_fp
# 
# hous_neighbor_fp
```

`r knit_child(hous_intro_fp)`

## Typology

```{r data-methods-typology}

# Setup scripts

make_typo_fp <- root_file('1-data/1-notebooks/04-make-typo-inds.Rmd')

# Visualization scripts

typology_fp <- root_file('1-data/1-notebooks/04-typology.Rmd')

```

`r knit_child(make_typo_fp)`

_This is a work in progress document - changes will be made._
<br>
`r knit_child(typology_fp)`

## Resources
_This is a work in progress document - changes will be made._
